---
title: "Advanced IV Estimation with ivreg2r"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced IV Estimation with ivreg2r}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette covers advanced features of `ivreg2r` beyond the basic 2SLS
workflow introduced in `vignette("introduction")`. Topics include:

- LIML, Fuller, and k-class estimation
- Instrument validity tests (orthogonality C-statistic)
- Weak-instrument-robust inference (Stock-Wright S statistic)
- Two-way clustering
- Reduced-form regression
- Weight types (overview)
- Degrees-of-freedom adjustments
- Factor variables

For HAC/AC standard errors, GMM, CUE, Kiefer, Driscoll-Kraay, partialling,
and user-supplied matrices, see `vignette("time-series-gmm")`.

We use three datasets: the Mroz (1987) female labor supply data, the
Griliches (1976) young men's wage data, and the Wooldridge wage panel.

```{r setup}
library(ivreg2r)
data(mroz)
data(griliches)
data(wagepan)
```

## The Mroz dataset

The Mroz (1987) dataset contains 753 married white women from the 1975 Panel
Study of Income Dynamics (PSID). Of these, 428 were working (`inlf == 1`) and
have observed wages. This is a classic IV application: years of education
(`educ`) is potentially endogenous in a wage equation due to ability bias.
Following the illustrative specification used throughout Stata's `ivreg2`
help file and in Baum, Schaffer & Stillman (2007, pp. 471--472), the excluded
instruments are `age` and the numbers of children at home --- under 6
(`kidslt6`) and 6--18 (`kidsge6`). These exclusion restrictions are
illustrative and contestable (children plausibly affect wage offers through
labor-supply channels); the help file uses the specification precisely
because it makes the instrument-testing machinery interesting.

```{r mroz-subset}
mroz_work <- subset(mroz, inlf == 1)
nrow(mroz_work)
```

Our baseline specification regresses log wages on experience and its square
(exogenous), with education instrumented by `age`, `kidslt6` (children under
6), and `kidsge6` (children 6-18):

```{r mroz-baseline}
fit_2sls <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                   data = mroz_work)
summary(fit_2sls)
```

This overidentified model (3 excluded instruments for 1 endogenous regressor,
hence 2 overidentifying restrictions) provides a foundation for comparing
estimators and exploring instrument diagnostics.

## LIML and Fuller estimation

### LIML

Limited Information Maximum Likelihood (LIML) is an alternative to 2SLS that
is more robust to weak instruments. Unlike 2SLS, LIML is approximately
median-unbiased and its Wald test suffers less size distortion when instruments
are weak (Stock & Yogo, 2005). However, LIML has no finite moments of any
order, regardless of the degree of overidentification (Mariano & Sawa, 1972),
so individual LIML estimates can be extreme in small samples; Fuller's
modification (below) restores finite moments. (It is 2SLS whose moments
depend on the degree of overidentification: the 2SLS estimator has moments
up to the number of overidentifying restrictions; Baum, Schaffer & Stillman,
2007, p. 490, fn. 19.) Use `method = "liml"`:

```{r liml}
fit_liml <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                   data = mroz_work, method = "liml")
summary(fit_liml)
```

The output reports `lambda`, the LIML eigenvalue. When `lambda` is close to 1,
LIML and 2SLS give similar results (2SLS is the special case `k = 1`). A
`lambda` far above 1 is evidence *against the overidentifying restrictions*,
not of weak instruments: `N * log(lambda)` is the Anderson-Rubin (1950)
overidentification LR statistic (see the Anderson-Rubin LIML
overidentification section below), so judge "how far above 1" by the reported
AR statistic and its p-value rather than by eye --- under weak but valid
instruments `lambda` stays near 1. Weak identification is diagnosed by the
Cragg-Donald and Kleibergen-Paap statistics, not by `lambda`.

### Fuller modification

Fuller's (1977) modified LIML uses `k = lambda - alpha / (N - L)` where `L`
is the number of instruments (including exogenous regressors) and `alpha` is
the Fuller parameter. Unlike LIML, Fuller's estimator has finite moments.
The standard choice is `fuller = 1`, suggested as a good choice by Fuller
(1977) and Davidson & MacKinnon (1993, pp. 649--650), as cited by Baum,
Schaffer & Stillman (2007, p. 479):

```{r fuller}
fit_fuller1 <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                      data = mroz_work, fuller = 1)
c(k = fit_fuller1$kclass_value, educ = unname(coef(fit_fuller1)["educ"]))
```

### k-class estimation

The k-class family nests OLS (`k = 0`), 2SLS (`k = 1`), and LIML
(`k = lambda`) as special cases. The `kclass` parameter accepts an arbitrary
`k`, but the documented non-trivial choice is Nagar's (1959) bias-adjusted
estimator with `k = 1 + (L - K) / N`, where `L - K` is the number of
overidentifying restrictions. Rather than hand-counting L and K, derive them
from the fitted baseline model --- `rankzz` is the rank of the full
instrument matrix (constant + 2 exogenous + 3 excluded = 6) and `rank` is the
number of estimated coefficients (constant + 2 exogenous + 1 endogenous = 4):

```{r kclass}
N <- nobs(fit_2sls)
L <- fit_2sls$rankzz
K <- fit_2sls$rank
k_nagar <- 1 + (L - K) / N
fit_nagar <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                    data = mroz_work, kclass = k_nagar)
c(L = L, K = K, k = k_nagar, educ = unname(coef(fit_nagar)["educ"]))
```

Research suggests that both Fuller's and Nagar's estimators have better
finite-sample performance than 2SLS in the presence of weak instruments,
although neither estimator is robust to violations of the i.i.d. assumption
(Baum, Schaffer & Stillman, 2007, pp. 478--479).

### When to use LIML

LIML is worth considering when the weak-identification diagnostics look
marginal: it has lower Stock-Yogo thresholds than 2SLS, meaning it tolerates
weaker instruments before inference becomes unreliable. Two qualifications
temper this into a recommendation rather than a rule. First, the Stock-Yogo
critical values are tabulated for the Cragg-Donald statistic under i.i.d.
errors; under a robust VCE they can be applied to the Kleibergen-Paap rk Wald
F only "with caution", or replaced by the Staiger-Stock F >= 10 rule of thumb
(Baum, Schaffer & Stillman, 2007, p. 490). Second, the ranking of LIML over
2SLS under weak instruments is itself established within the i.i.d.
Stock-Yogo framework (Baum, Schaffer & Stillman, 2007, pp. 489--490).
Switching estimators *because* a pretest flagged weakness also has the usual
pre-testing consequences for subsequent inference. When exactly identified
(one excluded instrument per endogenous variable), LIML and 2SLS are
numerically identical --- LIML's advantages only appear in the
overidentified case.

### COVIV

When using LIML or k-class with robust or cluster-robust standard errors, the
default sandwich bread uses the k-class projection matrix. The COVIV option
(`coviv = TRUE`) computes the covariance matrix using the 2SLS projection
(`k = 1`) rather than the LIML/k-class projection. Under iid errors, LIML with
COVIV is equivalent to CUE (see Estimator Equivalences below). With robust or
cluster-robust VCE, COVIV changes the bread matrix but uses the same LIML
residuals for the meat:

```{r coviv}
fit_coviv <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                    data = mroz_work, method = "liml", vcov = "robust",
                    small = TRUE, coviv = TRUE)
tidy(fit_coviv)[tidy(fit_coviv)$term == "educ", ]
```

### Comparing estimators

Use `tidy()` to build a comparison table across estimators:

```{r compare-estimators}
models <- list(
  "2SLS"      = fit_2sls,
  "LIML"      = fit_liml,
  "Fuller(1)" = fit_fuller1,
  "Nagar"     = fit_nagar
)
comparison <- do.call(rbind, lapply(names(models), function(nm) {
  tbl <- tidy(models[[nm]])
  tbl$model <- nm
  tbl
}))
# Show education coefficient across estimators
comparison[comparison$term == "educ", c("model", "estimate", "std.error")]
```

With strong instruments (as in this specification), all estimators produce
similar point estimates. Differences become more pronounced with weaker
instruments.

### Estimator equivalences

Several estimators coincide under specific conditions:

- **LIML + COVIV = CUE under conditional-homoskedasticity weighting**: when
  CUE uses the conditional-homoskedasticity (iid) weighting --- i.e.,
  `method = "cue"` with `vcov = "iid"` --- LIML and CUE produce
  mathematically identical coefficient estimates, and adding `coviv = TRUE`
  aligns the standard errors and test statistics as well. Because CUE is
  computed by numerical optimization, the agreement in practice is up to
  optimizer tolerance, not machine precision. CUE generalizes LIML to the
  non-iid case.
- **Driscoll-Kraay = cluster(time) + kernel**: DK VCE is equivalent
  to clustering on the time variable with a kernel (Bartlett by default) at
  the specified bandwidth. See `vignette("time-series-gmm")` for details.
- **Kiefer = truncated kernel at full bandwidth**: Kiefer VCE is equivalent
  to AC VCE with a truncated kernel where the bandwidth equals T (the number
  of time periods per panel unit). See `vignette("time-series-gmm")`.
- **LIML = 2SLS when exactly identified**: with one excluded instrument per
  endogenous variable, LIML reduces to 2SLS. LIML's advantages only appear
  in the overidentified case.

## Instrument validity tests

### Overidentification recap

With three excluded instruments and one endogenous regressor, the baseline
model has two overidentifying restrictions. The Sargan (1958) test is a joint
test of those restrictions --- a rejection signals that *some* instrument
fails the orthogonality conditions, but cannot attribute the failure to a
specific instrument:

```{r sargan}
glance(fit_2sls)[, c("overid_stat", "overid_p")]
```

### Orthogonality test (C-statistic)

The `orthog` argument tests whether a *subset* of instruments is exogenous,
conditional on the remaining instruments being valid. This implements the
difference-in-Sargan C-statistic (Hayashi, 2000).

The `ivreg2` help file demonstrates the test on the Griliches (1976) wage
equation, challenging the exogeneity of the two excluded instruments `age`
and `mrt` (marital status) while maintaining `med` (mother's education) and
`kww` (a test score). This replicates the help-file example at lines
1242--1243 (`orthog(age mrt)` with two-step GMM):

```{r orthog}
fit_orthog <- ivreg2(
  lw ~ s + expr + tenure + rns + smsa + factor(year) | iq |
    med + kww + age + mrt,
  data = griliches, method = "gmm2s", orthog = c("age", "mrt")
)
summary(fit_orthog)
```

The null hypothesis is that `age` and `mrt` satisfy the exclusion
restrictions, *given that the remaining instruments are valid*. Here the
C-statistic rejects emphatically: the data are inconsistent with treating
`age` and `mrt` as valid exclusions in this equation. (The same pair
reappears below as Baum, Schaffer & Stillman's weak-instruments
illustration.) The equation must remain identified after dropping the
challenged instruments, and with robust or cluster-robust standard errors
the test is computed as a difference in Hansen J statistics rather than
Sargan statistics.

### Anderson-Rubin LIML overidentification

When using LIML with iid errors, two additional overidentification statistics
are reported. The Anderson-Rubin (1950) LR statistic is `N * log(lambda)` and
its linearized form is `N * (lambda - 1)`, where `lambda` is the LIML
eigenvalue. Both are chi-squared with `L - K` degrees of freedom under the null
that the overidentifying restrictions hold. When `lambda` is close to 1, all
three overidentification tests (Sargan, AR LR, AR linearized) give similar
results:

```{r ar-overid}
glance(fit_liml)[, c("ar_overid_lr_stat", "ar_overid_lr_p",
                      "ar_overid_lin_stat", "ar_overid_lin_p")]
```

## Weak-instrument-robust inference

### Stock-Wright S statistic

The Stock-Wright (2000) S statistic is a weak-instrument-robust test of the
joint null hypothesis that the coefficients on the endogenous regressors are
zero and that the overidentifying restrictions hold. It tests the *same* null
as the Anderson-Rubin statistic reported in the same summary block, and the
two have the same distribution under that null (chi-squared with L1 degrees
of freedom): the Anderson-Rubin statistic is the Wald form of the test and
the S statistic is the LM / GMM-distance form (Baum, Schaffer & Stillman,
2007, pp. 491--492). Unlike Wald tests on 2SLS coefficients, both have
correct size even when instruments are weak. The S statistic is given by the
value of the CUE objective function (with exogenous regressors partialled
out), and is automatically computed for all IV models:

```{r stock-wright}
glance(fit_2sls)[, c("stock_wright_stat", "stock_wright_p")]
```

### When instruments are genuinely weak

Baum, Schaffer & Stillman (2007, pp. 492--493) illustrate the weak-instrument
problem with the Griliches (1976) wage equation, where `iq` (a measure of
ability) is instrumented by `age` and `mrt` (marital status). Their
specification includes year dummies and uses a heteroskedasticity-robust VCE
--- replicated faithfully here (the printed values match the published
output to all digits):

```{r weak-instruments}
fit_weak <- ivreg2(
  lw ~ s + expr + tenure + rns + smsa + factor(year) | iq | age + mrt,
  data = griliches, vcov = "robust"
)
summary(fit_weak)
```

Because the VCE is robust, the relevant weak-identification statistic is the
Kleibergen-Paap rk Wald F
(`r round(fit_weak$diagnostics$weak_id_robust$stat, 2)`), far below the
Stock-Yogo 10% maximal-size critical value of
`r fit_weak$diagnostics$weak_id_sy$critical_value[fit_weak$diagnostics$weak_id_sy$threshold == "10%"]`
--- bearing in mind BSS's own caveat that the Stock-Yogo values are
tabulated for the i.i.d. case and apply to the rk statistic only "with
caution" (Baum, Schaffer & Stillman, 2007, p. 490). With instruments this
weak, the Stock-Wright S statistic and the Anderson-Rubin test provide
size-correct inference even when 2SLS Wald tests do not. (BSS additionally
test whether `mrt` is a *redundant* instrument on this specification ---
the redundancy demonstration lives in `vignette("time-series-gmm")`.)

Two caveats on the robust tests themselves: they retain correct size but
lose power as instruments weaken (Baum, Schaffer & Stillman, 2007, p. 491),
and because the null is joint, a rejection can reflect either nonzero
coefficients on the endogenous regressors or failure of the overidentifying
restrictions --- the test cannot say which.

## Two-way clustering

### The wagepan dataset

The Wooldridge wage panel contains 4,360 observations on 545 young men from
the NLSY, observed annually from 1980 to 1987:

```{r wagepan-overview}
nrow(wagepan)
length(unique(wagepan$nr))    # persons
length(unique(wagepan$year))  # years
```

Panel data like this can exhibit correlation within individuals over time
*and* within years across individuals. Two-way clustering accounts for both.

### One-way vs. two-way clustering

Compare one-way clustering (by person) with two-way clustering (by person and
year):

```{r twoway}
fit_1way <- ivreg2(
  lwage ~ educ + black + hisp + exper + expersq + married + union,
  data = wagepan, clusters = ~ nr, small = TRUE
)

fit_2way <- ivreg2(
  lwage ~ educ + black + hisp + exper + expersq + married + union,
  data = wagepan, clusters = ~ nr + year, small = TRUE
)
```

Compare the standard errors:

```{r twoway-compare}
se_comparison <- data.frame(
  term = names(coef(fit_1way)),
  se_1way = tidy(fit_1way)$std.error,
  se_2way = tidy(fit_2way)$std.error
)
se_comparison$ratio <- se_comparison$se_2way / se_comparison$se_1way
se_comparison
```

Two-way clustering uses the Cameron, Gelbach & Miller (2011) formula:
`V_twoway = V_cluster1 + V_cluster2 - V_intersection`. The small-sample
correction uses `min(M1, M2)` clusters, and t-tests use
`min(M1, M2) - 1 = 7` degrees of freedom. With only 8 effective clusters,
users should be cautious: the CGM two-way cluster-robust VCE requires both
cluster dimensions to grow to infinity for consistency.

**Note:** Two-way clustering is only useful when the two cluster variables
define distinct, *crossed* (non-nested) categories --- as here, where every
person appears in several years and every year contains many persons. The
panel-by-time pairing is in fact the typical two-way application. If instead
one category is nested within the other (e.g., counties within states), the
two-way VCE reduces to the single-cluster VCE on the outer category, so
two-way clustering provides no additional robustness there.

## Reduced-form regression

The reduced form regresses the outcome directly on the full instrument set
(exogenous regressors plus excluded instruments), bypassing the first stage.

### Single-equation reduced form

Use `reduced_form = "rf"` to store the y-equation reduced form alongside the
structural estimates. With `vcov = "robust"`, this replicates the help-file
example at line 1385 (`robust ffirst saverf` on the Mroz baseline):

```{r rf}
fit_rf <- ivreg2(
  lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
  data = mroz_work, reduced_form = "rf", vcov = "robust"
)

# Reduced-form coefficients
fit_rf$reduced_form$coefficients

# The RF F-statistic for excluded instruments IS the Anderson-Rubin F ---
# the AR test is precisely this reduced-form F-test, which is why it
# remains valid under weak instruments.
fit_rf$reduced_form$f_stat
```

### System reduced form

Use `reduced_form = "system"` to estimate the full system --- one equation for
each endogenous variable plus the outcome --- with a cross-equation
variance-covariance matrix:

```{r rf-system}
fit_system <- ivreg2(
  lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
  data = mroz_work, reduced_form = "system"
)

# Equation names
fit_system$reduced_form$depvar

# Cross-equation VCV dimensions
dim(fit_system$reduced_form$vcov)
```

## Weight types

`ivreg2r` supports four weight types, matching Stata's `[aw=]`, `[fw=]`,
`[pw=]`, and `[iw=]` syntax. No worked weighted example exists in the
`ivreg2` help file or in the Baum-Schaffer-Stillman papers, so this section
documents semantics rather than fitting models; a worked sampling-weight
(pweight) example on the Card NLS data lives in `vignette("introduction")`.

- **Analytic weights** (`weight_type = "aweight"`, the default) are
  inverse-variance weights: observation i is treated as having error
  variance proportional to 1/w_i, the textbook case being data that arrive
  as group means computed from different numbers of underlying observations.
  Normalized internally to sum to N, so their scale is irrelevant.
- **Frequency weights** (`"fweight"`) must be positive integers: each row
  stands for `w` identical observations, and the effective sample size
  becomes `N = sum(weights)`, which enters every degrees-of-freedom
  calculation. Fitting on deduplicated data with fweights is identical to
  fitting on the fully expanded data.
- **Probability weights** (`"pweight"`) are sampling/survey weights
  (inverse selection probabilities). They force a robust VCE regardless of
  the `vcov` argument, because the weights describe the sampling design
  rather than the error variance. See the worked example in
  `vignette("introduction")`.
- **Importance weights** (`"iweight"`) have, in Stata's own description, no
  formal statistical definition --- each command defines how it treats them.
  In `ivreg2r` (as in `ivreg2`) they are not normalized, N becomes
  `sum(weights)` as a floating-point number, and they are restricted to the
  classical (iid) VCE --- no robust, cluster, kernel, or GMM combinations.

## Degrees-of-freedom adjustments

When fixed effects or other controls have been partialled out before
estimation, use `dofminus` and `sdofminus` to adjust the degrees of freedom
without re-estimating:

- `dofminus`: subtracted from N in large-sample variance formulas, so
  sigma^2 = RSS / (N - dofminus). The `ivreg2.ado` source describes it as
  the "large-sample adjustment (e.g., #fixed effects)"; it is how
  `xtivreg2, fe` reports correct statistics after within-demeaning consumes
  one degree of freedom per panel unit.
- `sdofminus`: subtracted from the residual degrees of freedom alongside K,
  entering the small-sample corrections --- per the same source comment,
  the "small-sample adjustment (e.g., #partialled-out regressors)", which
  is what Stata's `partial()` sets internally.

Neither option appears in the help file's examples, so we demonstrate
`sdofminus` with a self-verifying Frisch-Waugh-Lovell identity on the
canonical Griliches specification: residualize every variable on the year
dummies (7 columns including the intercept), refit without an intercept, and
tell `ivreg2()` that 7 degrees of freedom were consumed. The shared
coefficients, standard errors, and sigma must --- and do --- match the full
regression to machine precision:

```{r sdofminus-fwl}
fit_full <- ivreg2(
  lw ~ s + expr + tenure + rns + smsa + factor(year) | iq | med + kww + age,
  data = griliches, small = TRUE
)

# Partial the year dummies out of every variable by hand
yd <- model.matrix(~ factor(year), data = griliches)
resid_on_years <- function(v) lm.fit(yd, v)$residuals
gril_fwl <- data.frame(lapply(
  griliches[c("lw", "s", "expr", "tenure", "rns", "smsa",
              "iq", "med", "kww", "age")],
  resid_on_years
))

fit_fwl <- ivreg2(
  lw ~ 0 + s + expr + tenure + rns + smsa | iq | med + kww + age,
  data = gril_fwl, small = TRUE, sdofminus = 7
)

shared <- names(coef(fit_fwl))
max(abs(coef(fit_full)[shared] - coef(fit_fwl)[shared]))
max(abs(sqrt(diag(vcov(fit_full)))[shared] - sqrt(diag(vcov(fit_fwl)))[shared]))
c(sigma_full = fit_full$sigma, sigma_fwl = fit_fwl$sigma)
```

Without `sdofminus = 7`, the FWL coefficients would still match but the
standard errors and sigma would not: the refit would not know that 7
parameters were already spent on the year dummies. The built-in
`partial = "factor(year)"` argument performs exactly this
residualize-and-adjust sequence in one step (see
`vignette("time-series-gmm")` for the partialling workflow).

## Factor variables

`ivreg2r` handles factor (categorical) variables automatically, expanding
them into dummy variables using R's standard contrast coding. The
weak-instruments fit above already used one: `factor(year)` in the Griliches
specification plays the role of Stata's `xi i.year` / `_I*` year dummies in
the corresponding help-file and BSS examples. With treatment contrasts (the
default), one dummy per non-reference level appears in the output:

```{r factors}
coef(fit_weak)[grep("factor\\(year\\)", names(coef(fit_weak)))]
```

Prediction with `newdata` works correctly with factor variables --- the model
remembers the levels from estimation and applies the same contrasts, so
`newdata` only needs the regressor values (including a year observed in the
estimation sample):

```{r factor-predict}
new_obs <- data.frame(s = 12, expr = 5, tenure = 2, rns = 0, smsa = 1,
                      iq = 100, year = 67)
predict(fit_weak, newdata = new_obs)
```

## Model comparison workflow

A common workflow is to compare OLS, 2SLS, LIML, and Fuller(1) estimates for
the same specification. The `tidy()` and `glance()` methods make this easy:

```{r model-comparison}
models <- list(
  OLS       = ivreg2(lwage ~ exper + expersq + educ, data = mroz_work),
  `2SLS`    = fit_2sls,
  LIML      = fit_liml,
  `Fuller(1)` = fit_fuller1
)

# Coefficient comparison for education
coef_table <- do.call(rbind, lapply(names(models), function(nm) {
  tbl <- tidy(models[[nm]])
  tbl$model <- nm
  tbl
}))
coef_table[coef_table$term == "educ",
           c("model", "estimate", "std.error", "p.value")]
```

```{r model-diagnostics}
# Model-level diagnostics comparison
diag_table <- do.call(rbind, lapply(names(models), function(nm) {
  gl <- glance(models[[nm]])
  gl$model <- nm
  gl
}))
diag_table[, c("model", "method", "r.squared", "sigma", "nobs")]
```

## Publication tables with modelsummary

The [`modelsummary`](https://modelsummary.com/) package produces
publication-quality regression tables from any model with `tidy()` and
`glance()` methods. `ivreg2r` objects work out of the box.

```{r modelsummary-setup, eval = requireNamespace("modelsummary", quietly = TRUE)}
library(modelsummary)
```

### Basic table

```{r ms-basic, eval = requireNamespace("modelsummary", quietly = TRUE)}
modelsummary(fit_2sls, output = "markdown")
```

### Side-by-side comparison

Compare OLS, 2SLS, and LIML in a single table:

```{r ms-compare, eval = requireNamespace("modelsummary", quietly = TRUE)}
models <- list(
  "OLS"  = ivreg2(lwage ~ exper + expersq + educ, data = mroz_work),
  "2SLS" = fit_2sls,
  "LIML" = fit_liml
)
modelsummary(models, output = "markdown",
             gof_map = c("nobs", "r.squared", "adj.r.squared", "sigma"))
```

### Exponentiated coefficients

For log-linear models, `exponentiate = TRUE` reports multiplicative effects
(`exp(beta)`). `modelsummary` transforms the standard errors to the
exponentiated scale by the delta method (the displayed SE is
`exp(beta) * se(beta)`), so the table below is internally consistent on the
multiplicative scale:

```{r ms-exponentiate, eval = requireNamespace("modelsummary", quietly = TRUE)}
modelsummary(fit_2sls, output = "markdown", exponentiate = TRUE)
```

### Customizing output

Rename coefficients and select specific goodness-of-fit statistics:

```{r ms-custom, eval = requireNamespace("modelsummary", quietly = TRUE)}
modelsummary(
  list("2SLS" = fit_2sls, "LIML" = fit_liml),
  output = "markdown",
  coef_rename = c(educ = "Education", exper = "Experience",
                  expersq = "Experience²"),
  gof_map = c("nobs", "r.squared", "sigma", "weak_id_stat")
)
```

`modelsummary` supports many output formats including LaTeX, HTML, Word, and
gt tables. See `?modelsummary::modelsummary` for the full list.

## Stata migration

For a Stata-to-R translation table covering the most common options (basic
IV, robust/cluster, LIML/k-class, GMM, HAC, and more), see the Stata
migration section in `vignette("introduction")`; for the full argument list,
see `?ivreg2`.

## References

- Anderson, T.W. and Rubin, H. (1949). "Estimation of the Parameters of a
  Single Equation in a Complete System of Stochastic Equations." *Annals of
  Mathematical Statistics*, 20(1), 46--63.
- Anderson, T.W. and Rubin, H. (1950). "The Asymptotic Properties of
  Estimates of the Parameters of a Single Equation in a Complete System of
  Stochastic Equations." *Annals of Mathematical Statistics*, 21(4),
  570--582.
- Baum, C.F., Schaffer, M.E., and Stillman, S. (2007). "Enhanced Routines for
  Instrumental Variables/Generalized Method of Moments Estimation and
  Testing." *Stata Journal*, 7(4), 465--506.
- Cameron, A.C., Gelbach, J.B., and Miller, D.L. (2011). "Robust Inference
  with Multiway Clustering." *Journal of Business & Economic Statistics*,
  29(2), 238--249.
- Davidson, R. and MacKinnon, J.G. (1993). *Estimation and Inference in
  Econometrics*. Oxford University Press.
- Fuller, W.A. (1977). "Some Properties of a Modification of the Limited
  Information Estimator." *Econometrica*, 45(4), 939--953.
- Griliches, Z. (1976). "Wages of Very Young Men." *Journal of Political
  Economy*, 84(4), S69--S85.
- Hansen, L.P., Heaton, J., and Yaron, A. (1996). "Finite-Sample Properties
  of Some Alternative GMM Estimators." *Journal of Business & Economic
  Statistics*, 14(3), 262--280.
- Hayashi, F. (2000). *Econometrics*. Princeton University Press.
- Mariano, R.S. and Sawa, T. (1972). "The Exact Finite-Sample Distribution of
  the Limited-Information Maximum Likelihood Estimator in the Case of Two
  Included Endogenous Variables." *Journal of the American Statistical
  Association*, 67(337), 159--163.
- Mroz, T.A. (1987). "The Sensitivity of an Empirical Model of Married
  Women's Hours of Work to Economic and Statistical Assumptions."
  *Econometrica*, 55(4), 765--799.
- Nagar, A.L. (1959). "The Bias and Moment Matrix of the General k-Class
  Estimators of the Parameters in Simultaneous Equations." *Econometrica*,
  27(4), 575--595.
- Sargan, J.D. (1958). "The Estimation of Economic Relationships Using
  Instrumental Variables." *Econometrica*, 26(3), 393--415.
- Stock, J.H. and Wright, J.H. (2000). "GMM with Weak Identification."
  *Econometrica*, 68(5), 1055--1096.
- Stock, J.H. and Yogo, M. (2005). "Testing for Weak Instruments in Linear IV
  Regression." In D.W.K. Andrews and J.H. Stock (Eds.), *Identification and
  Inference for Econometric Models: Essays in Honor of Thomas Rothenberg*.
  Cambridge University Press.
