---
title: "Advanced IV Estimation with ivreg2r"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced IV Estimation with ivreg2r}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

This vignette covers advanced features of `ivreg2r` beyond the basic 2SLS
workflow introduced in `vignette("introduction")`. Topics include:

- LIML, Fuller, and k-class estimation
- Instrument validity tests (orthogonality C-statistic)
- Weak-instrument-robust inference (Stock-Wright S statistic)
- Two-way clustering
- Reduced-form regression
- Frequency and probability weights
- Degrees-of-freedom adjustments
- Factor variables

For HAC/AC standard errors, GMM, CUE, Kiefer, Driscoll-Kraay, partialling,
and user-supplied matrices, see `vignette("time-series-gmm")`.

We use two datasets: the Mroz (1987) female labor supply data and the
Wooldridge wage panel.

```{r setup}
library(ivreg2r)
data(mroz)
data(wagepan)
```

## The Mroz dataset

The Mroz (1987) dataset contains 753 married white women from the 1975 Panel
Study of Income Dynamics (PSID). Of these, 428 were working (`inlf == 1`) and
have observed wages. This is a classic IV application: years of education
(`educ`) is potentially endogenous in a wage equation due to ability bias, and
age and the number of young children at home serve as instruments. This is
the canonical specification used in Stata's `ivreg2` help file and originally
introduced by Wooldridge (2002).

```{r mroz-subset}
mroz_work <- subset(mroz, inlf == 1)
nrow(mroz_work)
```

Our baseline specification regresses log wages on experience and its square
(exogenous), with education instrumented by `age`, `kidslt6` (children under
6), and `kidsge6` (children 6-18):

```{r mroz-baseline}
fit_2sls <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                   data = mroz_work)
summary(fit_2sls)
```

This overidentified model (2 instruments for 1 endogenous variable) provides a
foundation for comparing estimators and exploring instrument diagnostics.

## LIML and Fuller estimation

### LIML

Limited Information Maximum Likelihood (LIML) is an alternative to 2SLS that
is more robust to weak instruments. Unlike 2SLS, LIML is approximately
median-unbiased and its Wald test suffers less size distortion when instruments
are weak (Stock & Yogo, 2005). However, LIML may lack finite moments,
especially when the degree of overidentification is small. Use
`method = "liml"`:

```{r liml}
fit_liml <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                   data = mroz_work, method = "liml")
summary(fit_liml)
```

The output reports `lambda`, the LIML eigenvalue. When `lambda` is close to 1,
LIML and 2SLS give similar results (2SLS is the special case `k = 1`). Large
values of `lambda` indicate potential weak instrument problems.

### Fuller modification

Fuller's (1977) modified LIML uses `k = lambda - alpha / (N - L)` where `L`
is the number of instruments (including exogenous regressors) and `alpha` is
the Fuller parameter. Unlike LIML, Fuller's estimator has finite moments.
The standard choice is `fuller = 1`, which Baum, Schaffer & Stillman (2007,
p. 479) cite as "a good choice" for approximate mean-unbiasedness:

```{r fuller}
fit_fuller1 <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                      data = mroz_work, fuller = 1)
```

### k-class estimation

The k-class family nests OLS (`k = 0`), 2SLS (`k = 1`), and LIML
(`k = lambda`) as special cases. The `kclass` parameter accepts an arbitrary
`k`, but the documented non-trivial choice is Nagar's (1959) bias-adjusted
estimator with `k = 1 + (L - K) / N`, where `L - K` is the number of
overidentifying restrictions:

```{r kclass}
N  <- nrow(mroz_work)
L  <- 5  # 2 exogenous + 1 const + 3 excluded instruments
K  <- 4  # 2 exogenous + 1 const + 1 endogenous
k_nagar <- 1 + (L - K) / N
fit_nagar <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                    data = mroz_work, kclass = k_nagar)
```

### When to use LIML

Use LIML when the Cragg-Donald or Kleibergen-Paap F statistic falls below the
Stock-Yogo critical values for 2SLS size distortion. LIML has lower Stock-Yogo
thresholds than 2SLS, meaning it tolerates weaker instruments before inference
becomes unreliable. When exactly identified (one excluded instrument per
endogenous variable), LIML and 2SLS are numerically identical --- LIML's
advantages only appear in the overidentified case.

### COVIV

When using LIML or k-class with robust or cluster-robust standard errors, the
default sandwich bread uses the k-class projection matrix. The COVIV option
(`coviv = TRUE`) computes the covariance matrix using the 2SLS projection
(`k = 1`) rather than the LIML/k-class projection. Under iid errors, LIML with
COVIV is equivalent to CUE (see Estimator Equivalences below). With robust or
cluster-robust VCE, COVIV changes the bread matrix but uses the same LIML
residuals for the meat:

```{r coviv}
fit_coviv <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                    data = mroz_work, method = "liml", vcov = "robust",
                    small = TRUE, coviv = TRUE)
```

### Comparing estimators

Use `tidy()` to build a comparison table across estimators:

```{r compare-estimators}
models <- list(
  "2SLS"      = fit_2sls,
  "LIML"      = fit_liml,
  "Fuller(1)" = fit_fuller1,
  "Nagar"     = fit_nagar
)
comparison <- do.call(rbind, lapply(names(models), function(nm) {
  tbl <- tidy(models[[nm]])
  tbl$model <- nm
  tbl
}))
# Show education coefficient across estimators
comparison[comparison$term == "educ", c("model", "estimate", "std.error")]
```

With strong instruments (as in this specification), all estimators produce
similar point estimates. Differences become more pronounced with weaker
instruments.

### Estimator equivalences

Several estimators coincide under specific conditions:

- **LIML + COVIV = CUE under iid errors**: under iid errors, LIML and CUE
  produce identical coefficient estimates. Adding `coviv = TRUE` further
  aligns the standard errors and test statistics, making the full output
  match CUE to machine precision. CUE generalizes LIML to the non-IID case.
- **Driscoll-Kraay = cluster(time) + kernel**: DK VCE is equivalent
  to clustering on the time variable with a kernel (Bartlett by default) at
  the specified bandwidth. See `vignette("time-series-gmm")` for details.
- **Kiefer = truncated kernel at full bandwidth**: Kiefer VCE is equivalent
  to AC VCE with a truncated kernel where the bandwidth equals T (the number
  of time periods per panel unit). See `vignette("time-series-gmm")`.
- **LIML = 2SLS when exactly identified**: with one excluded instrument per
  endogenous variable, LIML reduces to 2SLS. LIML's advantages only appear
  in the overidentified case.

## Instrument validity tests

### Overidentification recap

With two instruments and one endogenous variable, the Sargan (1958) test
checks whether both instruments satisfy the exclusion restriction:

```{r sargan}
glance(fit_2sls)[, c("overid_stat", "overid_p")]
```

### Orthogonality test (C-statistic)

The `orthog` argument tests whether a *subset* of instruments is exogenous,
conditional on the remaining instruments being valid. This implements the
difference-in-Sargan C-statistic (Hayashi, 2000).

For example, to test whether `kidslt6` is a valid instrument while
maintaining `age` and `kidsge6`:

```{r orthog}
fit_orthog <- ivreg2(
  lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
  data = mroz_work, orthog = c("kidslt6")
)
summary(fit_orthog)
```

The null hypothesis is that `kidslt6` satisfies the exclusion restriction,
*given that `age` and `kidsge6` are valid instruments*. Failure to reject is
consistent with its validity.

The test also works with robust and cluster-robust standard errors, where it
uses the Hansen J difference form:

```{r orthog-robust}
fit_orthog_robust <- ivreg2(
  lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
  data = mroz_work, orthog = c("kidslt6"), vcov = "robust", small = TRUE
)
glance(fit_orthog_robust)[, c("orthog_stat", "orthog_p")]
```

### Anderson-Rubin LIML overidentification

When using LIML with iid errors, two additional overidentification statistics
are reported. The Anderson-Rubin (1950) LR statistic is `N * log(lambda)` and
its linearized form is `N * (lambda - 1)`, where `lambda` is the LIML
eigenvalue. Both are chi-squared with `L - K` degrees of freedom under the null
that the overidentifying restrictions hold. When `lambda` is close to 1, all
three overidentification tests (Sargan, AR LR, AR linearized) give similar
results:

```{r ar-overid}
glance(fit_liml)[, c("ar_overid_lr_stat", "ar_overid_lr_p",
                      "ar_overid_lin_stat", "ar_overid_lin_p")]
```

## Weak-instrument-robust inference

### Stock-Wright S statistic

The Stock-Wright (2000) S statistic is a weak-instrument-robust test of the
joint null hypothesis that the coefficients on the endogenous regressors are
zero and that the overidentifying restrictions hold. Unlike Wald tests on 2SLS
coefficients, the S statistic has correct size even when instruments are weak.
It uses a GMM-distance formulation and is automatically computed for all IV
models:

```{r stock-wright}
glance(fit_2sls)[, c("stock_wright_stat", "stock_wright_p")]
```

Under iid errors with 2SLS, the S statistic and the Sargan statistic are
closely related --- both are based on the same GMM objective function, though
they test different null hypotheses (the S statistic includes the joint null
that the endogenous coefficients are zero).

### When instruments are genuinely weak

Baum, Schaffer & Stillman (2007, p. 492) illustrate the weak-instrument
problem with the Griliches (1976) wage equation, where `iq` (a measure of
ability) is instrumented by `age` and `mrt` (marital status). Both
instruments are weakly correlated with `iq` and the resulting Cragg-Donald
F statistic falls well below conventional thresholds:

```{r weak-instruments}
data(griliches)
fit_weak <- ivreg2(
  lw ~ s + expr + tenure + rns + smsa | iq | age + mrt,
  data = griliches
)
summary(fit_weak)
```

Note the weak identification F statistic --- with these marginal instruments,
the Cragg-Donald F drops well below Stock-Yogo critical values for 2SLS. In
such cases, the Stock-Wright S statistic and the Anderson-Rubin test provide
size-correct inference even when 2SLS Wald tests do not.

## Two-way clustering

### The wagepan dataset

The Wooldridge wage panel contains 4,360 observations on 545 young men from
the NLSY, observed annually from 1980 to 1987:

```{r wagepan-overview}
nrow(wagepan)
length(unique(wagepan$nr))    # persons
length(unique(wagepan$year))  # years
```

Panel data like this can exhibit correlation within individuals over time
*and* within years across individuals. Two-way clustering accounts for both.

### One-way vs. two-way clustering

Compare one-way clustering (by person) with two-way clustering (by person and
year):

```{r twoway}
fit_1way <- ivreg2(
  lwage ~ educ + black + hisp + exper + expersq + married + union,
  data = wagepan, clusters = ~ nr, small = TRUE
)

fit_2way <- ivreg2(
  lwage ~ educ + black + hisp + exper + expersq + married + union,
  data = wagepan, clusters = ~ nr + year, small = TRUE
)
```

Compare the standard errors:

```{r twoway-compare}
se_comparison <- data.frame(
  term = names(coef(fit_1way)),
  se_1way = tidy(fit_1way)$std.error,
  se_2way = tidy(fit_2way)$std.error
)
se_comparison$ratio <- se_comparison$se_2way / se_comparison$se_1way
se_comparison
```

Two-way clustering uses the Cameron, Gelbach & Miller (2011) formula:
`V_twoway = V_cluster1 + V_cluster2 - V_intersection`. The small-sample
correction uses `min(M1, M2)` clusters, and t-tests use
`min(M1, M2) - 1 = 7` degrees of freedom. With only 8 effective clusters,
users should be cautious: the CGM two-way cluster-robust VCE requires both
cluster dimensions to grow to infinity for consistency.

**Note:** Two-way clustering is only useful when the two cluster variables
define distinct, non-nested categories. If one category is nested within the
other (e.g., students within schools, or time periods within panels), the
two-way VCE reduces to the single-cluster VCE on the larger (outer) category,
so two-way clustering provides no additional robustness.

## Reduced-form regression

The reduced form regresses the outcome directly on the full instrument set
(exogenous regressors plus excluded instruments), bypassing the first stage.

### Single-equation reduced form

Use `reduced_form = "rf"` to store the y-equation reduced form alongside the
structural estimates:

```{r rf}
fit_rf <- ivreg2(
  lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
  data = mroz_work, reduced_form = "rf"
)

# Reduced-form coefficients
fit_rf$reduced_form$coefficients

# The RF F-statistic for excluded instruments IS the Anderson-Rubin F ---
# the AR test is precisely this reduced-form F-test, which is why it
# remains valid under weak instruments.
fit_rf$reduced_form$f_stat
```

### System reduced form

Use `reduced_form = "system"` to estimate the full system --- one equation for
each endogenous variable plus the outcome --- with a cross-equation
variance-covariance matrix:

```{r rf-system}
fit_system <- ivreg2(
  lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
  data = mroz_work, reduced_form = "system"
)

# Equation names
fit_system$reduced_form$depvar

# Cross-equation VCV dimensions
dim(fit_system$reduced_form$vcov)
```

## Weight types

`ivreg2r` supports three weight types, matching Stata's `[aw=]`, `[fw=]`, and
`[pw=]` syntax.

### Analytic weights (default)

Analytic weights (`weight_type = "aweight"`, the default) are the standard
weighted least squares weights, normalized internally to sum to N:

```{r aweight}
fit_aw <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                 data = mroz_work, weights = hours, weight_type = "aweight")
```

### Frequency weights

Frequency weights (`weight_type = "fweight"`) treat each observation as
representing `w` identical observations. The effective sample size becomes
`N = sum(weights)`, which affects all degrees-of-freedom calculations:

```{r fweight}
# Create an integer frequency weight for illustration
mroz_work$fw <- as.integer(pmin(mroz_work$hours %/% 500, 5) + 1L)

fit_fw <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                 data = mroz_work, weights = fw, weight_type = "fweight")
nobs(fit_fw)  # N = sum(fw), not nrow(mroz_work)
```

### Probability weights

Probability weights (`weight_type = "pweight"`) are sampling/survey weights.
They automatically force robust standard errors regardless of the `vcov`
argument:

```{r pweight}
fit_pw <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                 data = mroz_work, weights = hours, weight_type = "pweight")
fit_pw$vcov_type  # forced to "robust"
```

## Degrees-of-freedom adjustments

When fixed effects or other controls have been partialled out before
estimation, use `dofminus` and `sdofminus` to adjust the degrees of freedom
without re-estimating:

- `dofminus`: subtracted from N in large-sample variance formulas (affects
  sigma and VCV scaling)
- `sdofminus`: subtracted from residual df alongside K (affects small-sample
  corrections)

```{r dofminus}
fit_base <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                   data = mroz_work, small = TRUE)

fit_dof <- ivreg2(lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
                  data = mroz_work, small = TRUE,
                  dofminus = 5, sdofminus = 2)

# Compare sigma (RMSE)
c(base = fit_base$sigma, adjusted = fit_dof$sigma)

# Compare residual df
c(base = fit_base$df.residual, adjusted = fit_dof$df.residual)
```

This is useful when absorbing high-dimensional fixed effects: you partial them
out manually, then tell `ivreg2()` how many degrees of freedom were consumed.

## Factor variables

`ivreg2r` handles factor (categorical) variables automatically, expanding them
into dummy variables using R's standard contrast coding.

```{r factors}
# Create a three-level categorical variable from education
mroz_work$educ_cat <- cut(mroz_work$educ, breaks = c(0, 11, 13, 20),
                          labels = c("low", "mid", "high"))

fit_factor <- ivreg2(lwage ~ exper + expersq + educ_cat,
                     data = mroz_work)
coef(fit_factor)
```

The factor `educ_cat` has three levels; with treatment contrasts (the default),
two dummy variables appear in the output (`educ_catmid` and `educ_cathigh`),
with `low` as the reference category.

Prediction with `newdata` works correctly with factor variables --- the model
remembers the factor levels from estimation and applies the same contrasts:

```{r factor-predict}
new_obs <- data.frame(exper = 10, expersq = 100,
                      educ_cat = factor(c("low", "high"),
                                        levels = c("low", "mid", "high")))
predict(fit_factor, newdata = new_obs)
```

## Model comparison workflow

A common workflow is to compare OLS, 2SLS, LIML, and Fuller(1) estimates for
the same specification. The `tidy()` and `glance()` methods make this easy:

```{r model-comparison}
models <- list(
  OLS       = ivreg2(lwage ~ exper + expersq + educ, data = mroz_work),
  `2SLS`    = fit_2sls,
  LIML      = fit_liml,
  `Fuller(1)` = fit_fuller1
)

# Coefficient comparison for education
coef_table <- do.call(rbind, lapply(names(models), function(nm) {
  tbl <- tidy(models[[nm]])
  tbl$model <- nm
  tbl
}))
coef_table[coef_table$term == "educ",
           c("model", "estimate", "std.error", "p.value")]
```

```{r model-diagnostics}
# Model-level diagnostics comparison
diag_table <- do.call(rbind, lapply(names(models), function(nm) {
  gl <- glance(models[[nm]])
  gl$model <- nm
  gl
}))
diag_table[, c("model", "method", "r.squared", "sigma", "nobs")]
```

## Publication tables with modelsummary

The [`modelsummary`](https://modelsummary.com/) package produces
publication-quality regression tables from any model with `tidy()` and
`glance()` methods. `ivreg2r` objects work out of the box.

```{r modelsummary-setup, eval = requireNamespace("modelsummary", quietly = TRUE)}
library(modelsummary)
```

### Basic table

```{r ms-basic, eval = requireNamespace("modelsummary", quietly = TRUE)}
modelsummary(fit_2sls, output = "markdown")
```

### Side-by-side comparison

Compare OLS, 2SLS, and LIML in a single table:

```{r ms-compare, eval = requireNamespace("modelsummary", quietly = TRUE)}
models <- list(
  "OLS"  = ivreg2(lwage ~ exper + expersq + educ, data = mroz_work),
  "2SLS" = fit_2sls,
  "LIML" = fit_liml
)
modelsummary(models, output = "markdown",
             gof_map = c("nobs", "r.squared", "adj.r.squared", "sigma"))
```

### Exponentiated coefficients

For log-linear models, `exponentiate = TRUE` reports multiplicative effects
(`exp(beta)`). Standard errors remain on the log scale, following broom
convention:

```{r ms-exponentiate, eval = requireNamespace("modelsummary", quietly = TRUE)}
modelsummary(fit_2sls, output = "markdown", exponentiate = TRUE)
```

### Customizing output

Rename coefficients and select specific goodness-of-fit statistics:

```{r ms-custom, eval = requireNamespace("modelsummary", quietly = TRUE)}
modelsummary(
  list("2SLS" = fit_2sls, "LIML" = fit_liml),
  output = "markdown",
  coef_rename = c(educ = "Education", exper = "Experience",
                  expersq = "Experience²"),
  gof_map = c("nobs", "r.squared", "sigma", "weak_id_stat")
)
```

`modelsummary` supports many output formats including LaTeX, HTML, Word, and
gt tables. See `?modelsummary::modelsummary` for the full list.

## Stata migration

For a comprehensive Stata-to-R translation table covering all features (basic
IV, robust/cluster, LIML/k-class, GMM, HAC, and more), see the Stata migration
section in `vignette("introduction")`.

## References

- Anderson, T.W. and Rubin, H. (1949). "Estimation of the Parameters of a
  Single Equation in a Complete System of Stochastic Equations." *Annals of
  Mathematical Statistics*, 20(1), 46--63.
- Cameron, A.C., Gelbach, J.B., and Miller, D.L. (2011). "Robust Inference
  with Multiway Clustering." *Journal of Business & Economic Statistics*,
  29(2), 238--249.
- Fuller, W.A. (1977). "Some Properties of a Modification of the Limited
  Information Estimator." *Econometrica*, 45(4), 939--953.
- Hansen, L.P., Heaton, J., and Yaron, A. (1996). "Finite-Sample Properties
  of Some Alternative GMM Estimators." *Journal of Business & Economic
  Statistics*, 14(3), 262--280.
- Hayashi, F. (2000). *Econometrics*. Princeton University Press.
- Mroz, T.A. (1987). "The Sensitivity of an Empirical Model of Married
  Women's Hours of Work to Economic and Statistical Assumptions."
  *Econometrica*, 55(4), 765--799.
- Sargan, J.D. (1958). "The Estimation of Economic Relationships Using
  Instrumental Variables." *Econometrica*, 26(3), 393--415.
- Stock, J.H. and Wright, J.H. (2000). "GMM with Weak Identification."
  *Econometrica*, 68(5), 1055--1096.
- Stock, J.H. and Yogo, M. (2005). "Testing for Weak Instruments in Linear IV
  Regression." In D.W.K. Andrews and J.H. Stock (Eds.), *Identification and
  Inference for Econometric Models: Essays in Honor of Thomas Rothenberg*.
  Cambridge University Press.
- Wooldridge, J.M. (2010). *Econometric Analysis of Cross Section and Panel
  Data*, 2nd ed. MIT Press.
