---
title: "Instrumental Variables Estimation with ivreg2r"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Instrumental Variables Estimation with ivreg2r}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

`ivreg2r` provides comprehensive instrumental variables and GMM estimation with
automatic diagnostic tests. It supports 2SLS, LIML, Fuller, k-class, two-step
efficient GMM, and continuously-updated (CUE) estimators with classical,
robust, cluster-robust, HAC, and Driscoll-Kraay standard errors. It is an R
implementation inspired by Stata's `ivreg2` (Baum, Schaffer & Stillman),
designed for applied researchers --- especially those migrating from Stata who
expect rich diagnostics reported alongside their IV estimates.

This vignette covers the core workflow: OLS, 2SLS, diagnostics, robust and
cluster-robust inference, weights, and broom integration. For LIML, Fuller,
and k-class estimation, weak-instrument-robust inference, and model-comparison
workflows, see `vignette("advanced-iv")`. For HAC/AC standard errors, GMM,
CUE, partialling, and panel VCE options, see `vignette("time-series-gmm")`.

We'll use the Card (1995) dataset throughout this vignette. Card uses
geographic proximity to a four-year college as an instrument for years of
education in a log-wage equation, exploiting the idea that growing up near
a college reduces the cost of attending and therefore shifts educational
attainment. The exclusion restriction --- that proximity affects wages only
through schooling --- is an identifying *assumption*, not something the data
can fully verify: proximity could in principle be correlated with unobserved
determinants of wages, a concern Card (1995) discusses at length.

```{r setup}
library(ivreg2r)
data(card)
```

## OLS baseline

With a one-part formula, `ivreg2()` estimates by OLS. We use the control set
from Card (1995, Table 2), which is also the specification in Example 15.4 of
Wooldridge (2020): experience and its square, race,
urban residence (`smsa`), and region (current South plus the 1966 residence
indicators `smsa66` and `reg662`--`reg669`):

```{r ols}
fit_ols <- ivreg2(
  lwage ~ educ + exper + expersq + black + smsa + south + smsa66 +
    reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669,
  data = card
)
summary(fit_ols)
```

The point estimates are identical to `lm()`, and the model uses the same
estimation infrastructure as the IV models below, so switching between OLS and
2SLS requires only a formula change. The default *reporting* follows Stata's
conventions, however: z statistics rather than t statistics, standard errors
without the finite-sample N − K correction, and Root MSE computed as
sqrt(RSS/N). Setting `small = TRUE` reproduces `lm()`'s standard errors,
t statistics, and sigma exactly. (For *weighted* models, coefficients and
`small = TRUE` standard errors still match `lm()`, but sigma is defined
differently --- see Weighted estimation below.)

## Two-stage least squares

`ivreg2r` uses a three-part formula: `y ~ exogenous | endogenous | excluded_instruments`. Each variable appears exactly once, and exogenous
regressors are automatically included as instruments.

Here we instrument `educ` (endogenous) with `nearc4` (grew up near a 4-year
college):

```{r tsls}
fit_iv <- ivreg2(
  lwage ~ exper + expersq + black + smsa + south + smsa66 +
    reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669 |
    educ | nearc4,
  data = card
)
summary(fit_iv)
```

All `r format(nobs(fit_iv), big.mark = ",")` observations are used, and the
IV estimate of the return to schooling
(`r round(coef(fit_iv)["educ"], 4)`, versus
`r round(coef(fit_ols)["educ"], 4)` by OLS) matches the published
replications of this specification.

The three parts of the formula are:

| Part | Contents | Example |
|------|----------|---------|
| LHS | Dependent variable | `lwage` |
| 1st RHS | Exogenous regressors | `exper + expersq + black + smsa + south + smsa66 + reg662 + ... + reg669` |
| 2nd RHS | Endogenous regressors | `educ` |
| 3rd RHS | Excluded instruments | `nearc4` |

The full instrument set is the union of the exogenous regressors and the
excluded instruments. You never list the exogenous regressors twice.

## Understanding the diagnostics

The `summary()` output reports several diagnostic tests automatically. This
section walks through each one, using the values from the fit above. (All
numbers quoted in the text are computed from the fitted object, never typed
in by hand.)

```{r diag-values, include = FALSE}
d <- fit_iv$diagnostics
cd_f <- d$weak_id$stat
sy_iv_size <- d$weak_id_sy
sy10 <- sy_iv_size$critical_value[sy_iv_size$threshold == "10%"]
sy15 <- sy_iv_size$critical_value[sy_iv_size$threshold == "15%"]
td <- tidy(fit_iv)
educ_z_p <- td$p.value[td$term == "educ"]
```

### Weak identification

The fit above reports a **Cragg-Donald Wald F** of `r round(cd_f, 2)`. This
statistic tests whether the excluded instruments are *weak* --- that is, only
weakly correlated with the endogenous regressors. Weak instruments lead to
biased 2SLS estimates and unreliable standard errors.

**Stock-Yogo critical values** provide formal thresholds for the F statistic.
Two types exist:

- **IV size**: the critical value such that the *maximal* rejection rate of a
  nominal 5% Wald test on the endogenous regressors' coefficients is no more
  than 10%, 15%, 20%, or 25%. "Maximal" means the worst case over all possible
  true parameter values. For one endogenous variable and one instrument, the
  10% maximal-size critical value is `r sy10`.
- **IV relative bias**: the critical value such that the *maximum* relative
  bias of 2SLS (relative to OLS bias) is no more than 5%, 10%, 20%, or 30%.
  These are tabulated only when the number of overidentifying restrictions is
  at least two --- the moments of the 2SLS estimator needed to define the
  relative bias do not otherwise exist (Stock & Yogo, 2005; Baum, Schaffer &
  Stillman, 2007, p. 490). That is why neither this just-identified fit nor
  the overidentified fit below displays them: both have fewer than two
  overidentifying restrictions.

Here the Cragg-Donald F of `r round(cd_f, 2)` falls short of the 10%
maximal-size critical value of `r sy10` but clears the 15% value of
`r sy15` --- a genuinely borderline case. A 5% Wald test on `educ` could
reject as often as 15% of the time under the null, so the conventional
t-statistic should be read with some suspicion. This is exactly the situation
the weak-instrument-robust Anderson-Rubin test (below) is designed for.

Two caveats on using these critical values:

- The Stock-Yogo critical values are tabulated for the Cragg-Donald statistic
  under **i.i.d. errors** (the summary output prints a reminder when a robust
  VCE is in use). Under heteroskedasticity, autocorrelation, or clustering,
  the reported weak-identification statistic is the Kleibergen-Paap rk Wald F,
  and Baum, Schaffer & Stillman (2007, p. 490) recommend either applying the
  i.i.d. Stock-Yogo critical values to it "with caution" or falling back on
  the older Staiger & Stock (1997) rule of thumb that F should be at least 10.
- Under i.i.d. errors the Stock-Yogo tables give more precise guidance than
  the F > 10 rule, accounting for the number of endogenous regressors and
  instruments and distinguishing size distortion from relative bias. Under
  non-i.i.d. errors, where the tables do not formally apply, the F > 10 rule
  is a co-equal fallback rather than a strictly dominated heuristic.

### Underidentification

The fit above reports an Anderson LM statistic of
`r round(fit_iv$diagnostics$underid$stat, 2)`
(p = `r format.pval(fit_iv$diagnostics$underid$p, digits = 2, eps = 1e-4)`).
The underidentification test asks whether the excluded instruments have *any*
correlation with the endogenous regressors. Formally, the null hypothesis is
that the matrix of reduced-form coefficients on the excluded instruments has
rank K1 − 1, where K1 is the number of endogenous regressors --- i.e., that
the equation is underidentified. With a single endogenous regressor, as here,
this reduces to the intuitive statement that the instruments are irrelevant.
Rejection is good news: it means the model is identified.

With *several* endogenous regressors, note that the test fails to reject if
**any one** of them is unidentified --- it is a test of the whole system, not
of each regressor separately. (The Angrist-Pischke and Sanderson-Windmeijer
first-stage statistics, described below, provide per-regressor diagnostics.)

Under iid errors, this is the Anderson (1951) canonical correlations LM test.
Under heteroskedasticity or clustering, the Kleibergen-Paap (2006) rk LM
statistic is used instead.

Underidentification is distinct from weak identification. The underidentification
test asks "is there *any* correlation at all?" while the weak identification
test asks "is there *enough* correlation for reliable inference?" Here the
underidentification test rejects decisively even though the instrument is
borderline-weak by the Stock-Yogo standard --- a common pattern, and a
reminder that rejecting underidentification is not by itself evidence of
strong instruments.

### Overidentification

When there are more excluded instruments than endogenous regressors, the model
is overidentified and we can test whether the instruments are valid (i.e.,
uncorrelated with the structural error). Adding proximity to a *two*-year
college (`nearc2`) as a second instrument gives the specification of Computer
Exercise C5, Chapter 15, in Wooldridge (2020):

```{r overid}
fit_overid <- ivreg2(
  lwage ~ exper + expersq + black + smsa + south + smsa66 +
    reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669 |
    educ | nearc2 + nearc4,
  data = card
)
summary(fit_overid)
```

The Sargan (1958) test (for iid models) or Hansen (1982) J test (for
robust/cluster models) tests the null hypothesis that all instruments are
valid. Here the Sargan statistic is
`r round(fit_overid$diagnostics$overid$stat, 2)`
(p = `r round(fit_overid$diagnostics$overid$p, 4)`), so the single
overidentifying restriction is not rejected. A rejection would have indicated
that the instruments are not satisfying the required orthogonality conditions
--- either because an instrument is not truly exogenous, or because it is
incorrectly excluded from the regression. The test is a joint test of
instrument validity *and* correct model specification (Baum, Schaffer &
Stillman, 2003, pp. 16--17).

Important caveats:

- The test is only available when the model is overidentified (more excluded
  instruments than endogenous regressors). With exact identification, there are
  no testable restrictions.
- The test has low power when instruments are weak. More generally, failure to
  reject does not guarantee instrument validity --- the test has no power
  against alternatives under which all instruments remain consistent for the
  *same* (wrong) parameter value, as when each instrument's invalidity is
  proportional to its strength. It can only detect violations that move the
  instruments' implied estimates apart.

To test a *subset* of instruments while maintaining the rest, use the `orthog`
argument --- see `vignette("advanced-iv")` for details.

### Endogeneity test (C-statistic)

The endogeneity test (also called the C-statistic or difference-in-Sargan test)
asks whether the regressors treated as endogenous are actually endogenous. The
null hypothesis is that they are exogenous. For the just-identified fit above,
the statistic is `r round(fit_iv$diagnostics$endogeneity$stat, 2)`
(p = `r round(fit_iv$diagnostics$endogeneity$p, 4)`): we cannot reject that
`educ` is exogenous. In that case OLS may be preferred because it is more
efficient --- this fit is a live example of the trade-off, since the IV point
estimate is similar to OLS but far less precise.

Three caveats apply when acting on this test:

- It *maintains* instrument validity under both the null and the alternative.
  If the instruments are themselves invalid, the test is uninformative about
  endogeneity.
- Like the overidentification test, it has low power when instruments are
  weak --- precisely the situation flagged by the weak-identification
  diagnostics above.
- Using the test result to *choose* between OLS and IV is a pre-test:
  subsequent inference from the chosen estimator does not account for the
  selection step, and the size of post-pretest tests can be badly distorted
  (Guggenberger, 2010).

The test is implemented as a difference in Sargan statistics (under iid) or a
difference in Hansen J statistics (under robust/cluster VCE). See Hayashi
(2000, pp. 218--221 and 233--234) for the underlying theory.

**Note:** In Stata's `ivreg2`, the endogeneity test is only computed when
explicitly requested via the `endog()` option. `ivreg2r` computes it
automatically for convenience, testing all endogenous regressors by default.
Use the `endog` argument to test a specific subset (e.g.,
`endog = "educ"`).

### Anderson-Rubin test

The Anderson-Rubin (1949) test provides inference on the endogenous regressors
that remains valid even when instruments are weak. Unlike the standard Wald or
t-test on 2SLS coefficients --- which can be severely distorted by weak
instruments --- the AR test has correct size regardless of instrument strength.

The null hypothesis is that the coefficients on the endogenous regressors are
jointly zero *and* the overidentifying restrictions hold. When the model is
overidentified, rejection could indicate either nonzero coefficients or invalid
overidentifying restrictions (or both). In the just-identified case, the
overidentifying restrictions are trivially satisfied and the test is purely
about the coefficients.

This serves as a useful cross-check: if the standard t-test says an endogenous
regressor is significant but the AR test does not reject, weak instruments may
be distorting inference. In the running example the two tests agree but the
gap is visible: the conventional z-test on `educ` gives
p = `r round(educ_z_p, 4)`, while the
Anderson-Rubin chi-squared test gives
p = `r round(fit_iv$diagnostics$anderson_rubin$chi2_p, 4)` --- still a
rejection at the 5% level, but a weaker one, consistent with the borderline
first stage.

Robustness to weak instruments comes at a price: as instruments weaken, the
power of the AR test declines --- Baum, Schaffer & Stillman (2007, p. 491)
describe this as an intuitively appealing feature, since weak instruments
*should* make confident inference harder. In overidentified models the AR
chi-squared statistic has L1 degrees of freedom (L1 = number of excluded
instruments), so power also falls as instruments are added without
strengthening identification. More powerful weak-instrument-robust procedures
of the conditional likelihood ratio type exist for those cases (in Stata, see
`condivreg` and `rivtest`).

The Stock-Wright (2000) S statistic tests the same joint null hypothesis using
a GMM-distance formulation. It is also automatically computed. See
`vignette("advanced-iv")` for more on weak-instrument-robust inference.

## First-stage diagnostics

The summary output also reports a table of first-stage diagnostics for each
endogenous variable:

- **F-stat**: the partial F-test of the excluded instruments in the first-stage
  regression. This is the standard measure of instrument relevance for each
  endogenous variable individually. With multiple endogenous regressors, high
  individual F statistics are necessary but not sufficient --- the Cragg-Donald
  or Kleibergen-Paap multivariate statistics provide the proper joint test.
- **Partial R-sq**: the partial R-squared from the first-stage regression,
  measuring the fraction of variation in the endogenous variable explained by
  the excluded instruments after partialling out the exogenous regressors.
- **Shea partial R-sq**: Shea's (1997) partial R-squared, which takes the
  intercorrelations among the *instruments* into account. When
  the instruments predict multiple endogenous variables in similar ways, Shea's
  measure will be lower than the standard partial R-squared, signaling a
  potential identification problem. With a single endogenous variable, it
  equals the standard partial R-squared.
- **SW F** and **AP F**: Sanderson-Windmeijer (2016) and Angrist-Pischke
  (2009) conditional F statistics. With a single endogenous variable, SW F =
  AP F = the standard first-stage F. These become informative when there are
  multiple endogenous regressors, providing variable-by-variable diagnostics.

## Robust and clustered inference

### Heteroskedasticity-robust standard errors

Use `vcov = "robust"` for White (1980) heteroskedasticity-robust standard errors.
Add `small = TRUE` for a finite-sample correction (equivalent to Stata's `robust small`):

```{r robust}
fit_robust <- ivreg2(
  lwage ~ exper + expersq + black + smsa + south + smsa66 +
    reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669 |
    educ | nearc4,
  data = card, vcov = "robust", small = TRUE
)
summary(fit_robust)
```

Notice that the diagnostics automatically adapt. The underidentification test
is now the Kleibergen-Paap rk LM statistic instead of the Anderson LM
statistic. In the weak-identification block, the Kleibergen-Paap rk Wald F
(`r round(fit_robust$diagnostics$weak_id_robust$stat, 2)` here) is reported
*alongside* the Cragg-Donald Wald F (`r round(fit_robust$diagnostics$weak_id$stat, 2)`),
which remains in the output for reference --- only the rk statistic is valid
under non-i.i.d. errors. In overidentified models the Hansen J replaces the
Sargan statistic (this fit is exactly identified, so the line reads
"excluded").

The `small = TRUE` option applies finite-sample corrections throughout:
t-statistics and F-tests replace z-statistics and chi-squared tests.

### Cluster-robust standard errors

When observations are correlated within groups (e.g., repeated observations
on the same individual in panel data), use the `clusters` argument. The
`ivreg2` help file (line 1618) illustrates clustering with an **OLS**
regression --- no instruments are involved --- of log wages on grade, age,
total labor-market experience, and job tenure in the NLS young women's panel
(`nlswork`), clustering on the individual (`idcode`):

```{r cluster}
data(nlswork)
fit_cluster <- ivreg2(
  ln_wage ~ grade + age + ttl_exp + tenure,
  data = nlswork, clusters = ~ idcode
)
summary(fit_cluster)
```

```{r cluster-values, include = FALSE}
fit_iid <- ivreg2(ln_wage ~ grade + age + ttl_exp + tenure, data = nlswork)
se_cluster_grade <- sqrt(vcov(fit_cluster)["grade", "grade"])
se_iid_grade <- sqrt(vcov(fit_iid)["grade", "grade"])
```

We use this OLS example, taken directly from the help file, purely to
illustrate the `clusters` argument; the same option applies identically when
the formula has endogenous and excluded-instrument parts. Cluster-robust VCE
is automatically used whenever `clusters` is supplied, without needing to set
`vcov` explicitly. Fitting on the full
`r format(nobs(fit_cluster), big.mark = ",")` observations, the
cluster-robust standard error on `grade` is
`r formatC(se_cluster_grade, format = "f", digits = 4)`, versus
`r formatC(se_iid_grade, format = "f", digits = 4)` under the
classical (i.i.d.) VCE fit on the same data --- clustering roughly doubles
the SE here because wages are serially correlated within person across the
panel's multiple survey years, a dependence that classical and
heteroskedasticity-robust SEs both ignore.

Cluster-robust inference is asymptotic in the number of clusters, not the
number of observations: with only a handful of clusters (e.g., a dozen
states or regions), the standard errors can be badly biased downward even
when N is large, and the `small = TRUE` finite-sample correction only
partially compensates. Here there are
`r format(fit_cluster$n_clusters, big.mark = ",")` clusters, one per
individual, so this concern does not arise; researchers clustering on a
small number of groups should treat the reported SEs with caution and
consider wild-cluster bootstrap or other few-cluster methods.

When `small = TRUE` is added, `ivreg2r` applies the standard cluster
finite-sample correction factor, `(N-1)/(N-K) * M/(M-1)`, where M is the
number of clusters, and reports t- and F-statistics instead of z- and
chi-squared statistics.

For two-way clustering (e.g., by firm and year in a panel), use
`clusters = ~ var1 + var2`. See `vignette("advanced-iv")` for details.

### VCE option summary

| `vcov` | `clusters` | `small` | Equivalent Stata command |
|--------|-----------|---------|------------------------|
| `"iid"` | — | `FALSE` | `ivreg2 ...` |
| `"iid"` | — | `TRUE` | `ivreg2 ..., small` |
| `"robust"` | — | `FALSE` | `ivreg2 ..., robust` |
| `"robust"` | — | `TRUE` | `ivreg2 ..., robust small` |
| — | `~ var` | `FALSE` | `ivreg2 ..., cluster(var)` |
| — | `~ var` | `TRUE` | `ivreg2 ..., cluster(var) small` |

When `clusters` is supplied, VCE is automatically cluster-robust regardless of
the `vcov` argument.

For HAC, AC, Kiefer, and Driscoll-Kraay VCE options, see
`vignette("time-series-gmm")`.

## Weighted estimation

The Card data include a genuine survey weight: `weight` is the NLS sampling
weight, reflecting each respondent's probability of selection into the survey.
Probability/sampling weights (equivalent to Stata's `[pw=varname]`) are
specified with `weights` plus `weight_type = "pweight"`:

```{r weights}
fit_pw <- ivreg2(
  lwage ~ exper + expersq + black + smsa + south + smsa66 +
    reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669 |
    educ | nearc4,
  data = card, weights = weight, weight_type = "pweight"
)
summary(fit_pw)
```

Because pweights describe the sampling design rather than the error variance,
valid standard errors must be robust: `ivreg2r` forces a robust VCE whenever
`weight_type = "pweight"`, regardless of the `vcov` setting (the message above
notes the override), exactly as Stata does for `[pw=]`. The weighted point
estimate answers a population-level question --- what is the return to
schooling in the population the NLS sample was drawn from --- whereas the
unweighted fit describes the sample.

`ivreg2r` also supports analytic weights (the default `weight_type`,
equivalent to Stata's `[aw=]`), frequency weights (`"fweight"`), and
importance weights (`"iweight"`). For aweights and pweights, the weights are
normalized internally to sum to N (Stata's convention), so coefficients,
standard errors, and all test statistics are unaffected by the scale of the
weights; one consequence is that for these types sigma (the Root MSE)
differs from `lm()`'s by a factor of `sqrt(N / sum(w))`, because `lm()` uses
raw weights. Frequency and importance weights are *not* normalized --- they
redefine the sample size as `sum(weights)`, so for them the scale of the
weights matters by construction. See `?ivreg2` for details on each weight
type.

## Beyond 2SLS

This vignette focuses on 2SLS, but `ivreg2r` also supports LIML, Fuller,
and k-class estimation via the `method`, `fuller`, and `kclass` arguments.
LIML has smaller approximate bias than 2SLS when there are many or weak
instruments. Fuller's modified LIML further improves on LIML by having finite
moments. These finite-sample advantages are established under i.i.d. errors;
neither estimator is robust to violations of that assumption (Baum, Schaffer
& Stillman, 2007, pp. 478--479). See `vignette("advanced-iv")` for examples
and comparisons.

## Confidence intervals

`confint()` returns confidence intervals for all coefficients:

```{r confint}
confint(fit_iv)
```

When `small = TRUE` was used at estimation time, intervals are based on the
t-distribution; otherwise, the standard normal.

## Tidyverse workflows

`ivreg2` objects integrate with the tidyverse through `tidy()`, `glance()`,
and `augment()`.

### Coefficient table

`tidy()` returns a tibble of coefficient estimates with standard errors,
test statistics, p-values, and confidence intervals:

```{r tidy}
tidy(fit_iv)
```

### Model summary

`glance()` returns a single-row tibble with fit statistics and all diagnostic
test results --- useful for building model comparison tables:

```{r glance}
glance(fit_iv)
```

### Augmented data

`augment()` returns the original data with fitted values and residuals appended:

```{r augment}
augment(fit_iv) |> head()
```

## Stata migration quick reference

| Task | Stata | R (`ivreg2r`) |
|------|-------|---------------|
| Basic IV | `ivreg2 y x (endo = z)` | `ivreg2(y ~ x | endo | z, data = d)` |
| No endogenous regressors (HOLS form) | `ivreg2 y x (=z1 z2)` | `ivreg2(y ~ x | 0 | z1 + z2, data = d)` |
| Robust SEs | `ivreg2 ..., robust` | `ivreg2(..., vcov = "robust")` |
| Robust + small | `ivreg2 ..., robust small` | `ivreg2(..., vcov = "robust", small = TRUE)` |
| Cluster SEs | `ivreg2 ..., cluster(g)` | `ivreg2(..., clusters = ~ g)` |
| Cluster + small | `ivreg2 ..., cluster(g) small` | `ivreg2(..., clusters = ~ g, small = TRUE)` |
| Analytic weights | `ivreg2 ... [aw=w]` | `ivreg2(..., weights = w)` |
| Coefficients | `e(b)` | `coef(fit)` |
| VCV matrix | `e(V)` | `vcov(fit)` |
| Residuals | `predict resid, resid` | `residuals(fit)` |
| Fitted values | `predict yhat` | `fitted(fit)` |
| Confidence intervals | — | `confint(fit)` |
| Tidy coef table | — | `tidy(fit)` |
| Summary statistics | `ereturn list` | `glance(fit)` |
| Diagnostics | Displayed after estimation | Displayed by `summary(fit)` |
| LIML | `ivreg2 ..., liml` | `ivreg2(..., method = "liml")` |
| Fuller(1) | `ivreg2 ..., fuller(1)` | `ivreg2(..., fuller = 1)` |
| k-class | `ivreg2 ..., kclass(0.5)` | `ivreg2(..., kclass = 0.5)` |
| COVIV | `ivreg2 ..., liml coviv` | `ivreg2(..., method = "liml", coviv = TRUE)` |
| Two-way cluster | `ivreg2 ..., cluster(g1 g2)` | `ivreg2(..., clusters = ~ g1 + g2)` |
| Orthogonality test | `ivreg2 ..., orthog(z1)` | `ivreg2(..., orthog = c("z1"))` |
| Frequency weights | `ivreg2 ... [fw=w]` | `ivreg2(..., weights = w, weight_type = "fweight")` |
| Probability weights | `ivreg2 ... [pw=w]` | `ivreg2(..., weights = w, weight_type = "pweight")` |
| DoF adjustment | `ivreg2 ..., dofminus(5)` | `ivreg2(..., dofminus = 5)` |
| Save reduced form | `ivreg2 ..., saverf` | `ivreg2(..., reduced_form = "rf")` (Stata's `rf` *displays* the reduced form; `saverf` *stores* it, which is what `reduced_form = "rf"` does) |
| GMM2S | `ivreg2 ..., gmm2s` | `ivreg2(..., method = "gmm2s")` |
| CUE | `ivreg2 ..., cue` | `ivreg2(..., method = "cue")` |
| AC kernel | `ivreg2 ..., kernel(bar) bw(3)` | `ivreg2(..., kernel = "bartlett", bw = 3, tvar = "year")` |
| HAC kernel | `ivreg2 ..., robust kernel(bar) bw(3)` | `ivreg2(..., vcov = "robust", kernel = "bartlett", bw = 3, tvar = "year")` |
| Auto bandwidth | `ivreg2 ..., kernel(bar) bw(auto)` | `ivreg2(..., kernel = "bartlett", bw = "auto", tvar = "year")` |
| Driscoll-Kraay | `ivreg2 ..., dkraay(3)` | `ivreg2(..., dkraay = 3, tvar = "year", ivar = "id")` |
| Kiefer VCE | `ivreg2 ..., kiefer` | `ivreg2(..., kiefer = TRUE, tvar = "year", ivar = "id")` |
| Partialling | `ivreg2 ..., partial(x1)` | `ivreg2(..., partial = c("x1"))` |
| Redundancy test | `ivreg2 ..., redundant(z1)` | `ivreg2(..., redundant = c("z1"))` |
| User W matrix | `ivreg2 ..., wmatrix(name)` | `ivreg2(..., wmatrix = W)` |
| Center moments | `ivreg2 ..., center` | `ivreg2(..., center = TRUE)` |

See `vignette("advanced-iv")` for LIML, k-class, two-way clustering, and model
comparison examples. See `vignette("time-series-gmm")` for HAC/AC, GMM, CUE,
partialling, and panel VCE examples.

### Intentional differences from Stata

| Area | Stata `ivreg2` | `ivreg2r` | Rationale |
|------|----------------|-----------|-----------|
| Formula | `ivreg2 y x (endo = z)` | `y ~ x \| endo \| z` | R convention; each variable listed once |
| Numerical method | Cross-products via Mata | QR-based regression via `lm.fit` | Better conditioning for core estimation |
| Diagnostics | Some computed at post-estimation | All computed at estimation time | Always available in `glance()` |
| Factor variables | `i.` prefix syntax | R `model.matrix()` / contrasts | R convention |
| COVIV | Implicit for LIML+robust | Explicit `coviv = TRUE` | Explicit > implicit |
| Time-series | `tsset` declares time/panel | `tvar=` / `ivar=` arguments | R has no `tsset`; explicit args |
| CUE optimizer | Mata `optimize()` | R `optim()` (BFGS + Nelder-Mead) | Different optimizer, same objective |

## References

- Anderson, T.W. (1951). "Estimating Linear Restrictions on Regression
  Coefficients for Multivariate Normal Distributions." *Annals of Mathematical
  Statistics*, 22(3), 327--351.
- Anderson, T.W. and Rubin, H. (1949). "Estimation of the Parameters of a
  Single Equation in a Complete System of Stochastic Equations." *Annals of
  Mathematical Statistics*, 20(1), 46--63.
- Angrist, J.D. and Pischke, J.-S. (2009). *Mostly Harmless Econometrics*.
  Princeton University Press.
- Baum, C.F., Schaffer, M.E., and Stillman, S. (2003). "Instrumental Variables
  and GMM: Estimation and Testing." *Stata Journal*, 3(1), 1--31.
- Baum, C.F., Schaffer, M.E., and Stillman, S. (2007). "Enhanced Routines for
  Instrumental Variables/Generalized Method of Moments Estimation and Testing."
  *Stata Journal*, 7(4), 465--506.
- Card, D. (1995). "Using Geographic Variation in College Proximity to Estimate
  the Return to Schooling." In L.N. Christofides, E.K. Grant, and R. Swidinsky
  (Eds.), *Aspects of Labour Market Behaviour: Essays in Honour of John
  Vanderkamp*. University of Toronto Press.
- Guggenberger, P. (2010). "The Impact of a Hausman Pretest on the Asymptotic
  Size of a Hypothesis Test." *Econometric Theory*, 26(2), 369--382.
- Hansen, L.P. (1982). "Large Sample Properties of Generalized Method of
  Moments Estimators." *Econometrica*, 50(4), 1029--1054.
- Hayashi, F. (2000). *Econometrics*. Princeton University Press.
- Kleibergen, F. and Paap, R. (2006). "Generalized Reduced Rank Tests Using
  the Singular Value Decomposition." *Journal of Econometrics*, 133(1),
  97--126.
- Sanderson, E. and Windmeijer, F. (2016). "A Weak Instrument F-Test in Linear
  IV Models with Multiple Endogenous Variables." *Journal of Econometrics*,
  190(2), 212--221.
- Sargan, J.D. (1958). "The Estimation of Economic Relationships Using
  Instrumental Variables." *Econometrica*, 26(3), 393--415.
- Shea, J. (1997). "Instrument Relevance in Multivariate Linear Models: A
  Simple Measure." *Review of Economics and Statistics*, 79(2), 348--352.
- Staiger, D. and Stock, J.H. (1997). "Instrumental Variables Regression with
  Weak Instruments." *Econometrica*, 65(3), 557--586.
- Stock, J.H. and Wright, J.H. (2000). "GMM with Weak Identification."
  *Econometrica*, 68(5), 1055--1096.
- Stock, J.H. and Yogo, M. (2005). "Testing for Weak Instruments in Linear IV
  Regression." In D.W.K. Andrews and J.H. Stock (Eds.), *Identification and
  Inference for Econometric Models: Essays in Honor of Thomas Rothenberg*.
  Cambridge University Press.
- White, H. (1980). "A Heteroskedasticity-Consistent Covariance Matrix
  Estimator and a Direct Test for Heteroskedasticity." *Econometrica*, 48(4),
  817--838.
- Wooldridge, J.M. (2020). *Introductory Econometrics: A Modern Approach*,
  7th edition. Cengage.
