---
title: "Time Series IV, Panel VCE, and GMM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time Series IV, Panel VCE, and GMM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

Heteroskedasticity-robust standard errors remain consistent under arbitrary
heteroskedasticity, but neither they nor classical standard errors survive
*autocorrelation* or *cross-sectional dependence* --- the settings this
vignette is about. It covers the tools `ivreg2r` provides for them:

- **HAC and AC standard errors** for time-series data
- **Panel VCE options**: Kiefer, Driscoll-Kraay, and cluster+kernel
- **Two-step efficient GMM** for overidentified models with non-IID errors
- **CUE estimation** as the GMM generalization of LIML
- **User-supplied W and S matrices** for iterated GMM and custom weighting
- **Partialling out regressors** for high-dimensional fixed effects
- **Additional options**: redundancy tests, moment centering, PSD corrections

We use five datasets: the U.S. Phillips-curve data from Wooldridge (annual,
1948--1996; pure time series), the Stock-Watson quarterly macro data used by
Baum, Schaffer & Stillman (2007), the Mroz (1987) labor-supply data, the
Card (1995) schooling data, and the Wooldridge wage panel.

```{r setup}
library(ivreg2r)
data(phillips)
data(stockwatson)
data(mroz)
data(card)
data(wagepan)
```

## HAC and AC standard errors

### The Phillips curve

The Phillips dataset contains 49 annual U.S. observations (1948--1996) on
inflation and unemployment. A simple expectations-augmented Phillips curve
regresses the change in inflation on unemployment:

```{r phillips-ols}
fit_iid <- ivreg2(cinf ~ unem, data = phillips)
```

Note that we pass the data frame as-is: `ivreg2()` drops incomplete cases
per specification, exactly as Stata does, so each fit uses the largest
estimation sample its own variables allow. (Running `na.omit()` on the whole
data frame first would also drop rows whose only missing values are in
*unused* lag columns --- here it would discard 1949 and silently change the
canonical results below.)

With time-series data, the errors are likely autocorrelated, making classical
SEs inconsistent. HAC (heteroskedasticity and autocorrelation consistent)
standard errors account for both.

### Specifying kernel-based standard errors

Use `kernel`, `bw` (bandwidth), and `tvar` to request kernel-based inference.
The type of VCE depends on the `vcov` setting: with `vcov = "robust"` (or
when heteroskedasticity-robust VCE is already active), you get **HAC**
(heteroskedasticity and autocorrelation consistent) standard errors. With the
default `vcov = "iid"`, specifying a kernel gives **AC** (autocorrelation
consistent) standard errors, which assume homoskedasticity but allow for
serial correlation. This matches Stata's `ivreg2`, where `bw`/`kernel` alone
produces AC and combining `robust` with `bw` produces HAC.

```{r hac-bartlett}
fit_hac <- ivreg2(cinf ~ unem, data = phillips,
                  vcov = "robust", kernel = "bartlett", bw = 3, tvar = "year",
                  small = TRUE)
summary(fit_hac)
```

This is the canonical HAC specification from Stata's `ivreg2` help file
(line 1511), reproduced on the full N = `r nobs(fit_hac)` estimation sample.
The Bartlett kernel applies linearly declining weights to autocovariances up
to lag `bw - 1`; combined with `vcov = "robust"`, this is the Newey-West
(1987) HAC estimator (the help file pairs the example with
`newey cinf unem, lag(2)`).

### Choosing a kernel

`ivreg2r` supports eight kernels, matching Stata's `ivreg2`:

| Kernel | Weights | Guaranteed PSD? | Typical use |
|--------|---------|-----------------|-------------|
| Bartlett | Linear decline | Yes | Default choice; Newey-West HAC |
| Parzen | Piecewise cubic | Yes | Smoother than Bartlett |
| Quadratic Spectral | $\sin/\cos$ based | Yes | Andrews (1991) |
| Truncated | Uniform up to lag | No | All lags equally weighted; Kiefer equivalence |
| Tukey-Hanning | Cosine taper | No | Standard spectral window |
| Tukey-Hamming | Cosine taper | No | Variant of Hanning |
| Daniell | $\sin(x)/x$ | No | Spectral analysis |
| Tent | Tent-shaped spectral window | No | Spectral analysis |

Bartlett, Parzen, and Quadratic Spectral guarantee a positive semi-definite
covariance estimate (Andrews 1991). For the other kernels, use the `psd`
option if negative eigenvalues arise.

### Automatic bandwidth

Use `bw = "auto"` for the Newey-West (1994) plug-in bandwidth selector, which
chooses the bandwidth from the data. This is available for Bartlett, Parzen,
and Quadratic Spectral kernels. Pairing it with the Quadratic Spectral
kernel replicates the help-file example at line 1504
(`ivreg2 cinf unem, kernel(qs) bw(auto)`):

```{r hac-auto}
fit_auto <- ivreg2(cinf ~ unem, data = phillips,
                   kernel = "qs", bw = "auto", tvar = "year")
fit_auto$bw  # selected bandwidth
```

### HAC vs AC

AC (autocorrelation consistent) standard errors assume homoskedasticity but
allow for autocorrelation. Use `vcov = "AC"` explicitly --- this is an exact
replication of the help-file AC example at line 1501
(`ivreg2 cinf unem, bw(3)`; Bartlett is Stata's default kernel):

```{r ac}
fit_ac <- ivreg2(cinf ~ unem, data = phillips,
                 vcov = "AC", kernel = "bartlett", bw = 3, tvar = "year")
```

HAC allows both heteroskedasticity and autocorrelation; AC assumes
homoskedastic errors with autocorrelation. When homoskedasticity holds, the
AC estimator may perform better in finite samples, at the cost of
inconsistency if it does not.

### Kernel-based VCE with IV

Kernel-based standard errors work with all estimation methods. Here we
instrument unemployment with its second lag, using HAC (adding
`vcov = "robust"`):

```{r hac-iv}
fit_iv_hac <- ivreg2(
  cinf ~ unem_1 | unem | unem_2,
  data = phillips, vcov = "robust", kernel = "bartlett", bw = 3, tvar = "year"
)
summary(fit_iv_hac)
```

The reported diagnostics automatically adapt to the kernel-based VCE: the
identification tests become Kleibergen-Paap rk statistics and the
endogeneity test becomes a difference in Hansen J statistics. (This
particular model is exactly identified --- one excluded instrument for one
endogenous regressor --- so the overidentification line reads "excluded";
the overidentified Stock-Watson example below shows a kernel-based Hansen
J in action.) Bear in mind that the Stock-Yogo critical values printed
alongside the weak-identification statistic are tabulated for the
Cragg-Donald statistic under i.i.d. errors; Baum, Schaffer & Stillman
(2007, p. 490) advise applying them to the Kleibergen-Paap rk Wald F only
with caution.

> **Stata parity note:** The Kleibergen-Paap identification tests
> (underidentification and weak identification) always use the Bartlett
> kernel internally, regardless of the kernel specified for the VCE. This
> matches the behavior of Stata's `ivreg2` --- not by design but by a quirk
> in its plumbing: `ranktest` itself accepts a `kernel()` option
> (ranktest.ado line 104), but `ivreg2.ado` captures only the bandwidth
> (`s(bwopt)`) from its option parser and never the kernel (`s(kernopt)`),
> so `ranktest` falls back to its Bartlett default. The overidentification
> test uses the user-specified kernel as expected.

## Panel VCE options

Panel data introduces richer correlation structures. Beyond one-way and
two-way clustering (covered in `vignette("introduction")` and
`vignette("advanced-iv")`), `ivreg2r` provides three kernel-based panel VCE
options.

### Kiefer VCE

Kiefer (1980) VCE assumes conditional homoskedasticity across all observations
but allows arbitrary intra-group (within-panel) serial correlation. It is
equivalent to AC VCE with a truncated kernel at bandwidth T − 1, which puts
weight one on every lag available within a panel of length T (bandwidths
beyond T − 1 are rejected, as in Stata):

```{r kiefer}
fit_kiefer <- ivreg2(
  lwage ~ educ + black + hisp + exper + expersq + married + union,
  data = wagepan, kiefer = TRUE, tvar = "year", ivar = "nr",
  small = TRUE
)
```

Use Kiefer when conditional homoskedasticity is plausible but you want to
allow for arbitrary within-unit serial correlation. If heteroskedasticity is
also a concern, use Driscoll-Kraay or cluster+kernel instead.

### Driscoll-Kraay VCE

Driscoll and Kraay (1998) VCE handles cross-sectional dependence *and*
persistence. It clusters on the time variable and applies a kernel to scores
summed within each time period:

```{r dk}
fit_dk <- ivreg2(
  lwage ~ educ + black + hisp + exper + expersq + married + union,
  data = wagepan, dkraay = 3, tvar = "year", ivar = "nr",
  small = TRUE
)
```

The `dkraay` argument specifies the bandwidth (here, Bartlett with bandwidth
3, i.e. autocovariances up to 2 lags --- for the Bartlett kernel the number
of lags is the bandwidth minus one). DK is robust to arbitrary
cross-sectional dependence --- useful when common
shocks (business cycles, policy changes) affect all individuals
simultaneously. DK is equivalent to clustering on the time variable with a
kernel (Bartlett by default; can be overridden with the `kernel` argument).

**Important:** Driscoll-Kraay is a large-T estimator --- the panel should have
a reasonably long time-series dimension for the asymptotics to apply. It is
not appropriate for panels with very few time periods.

### Cluster + kernel

Two-way cluster+kernel VCE accounts for both within-panel correlation (via
clustering on `ivar`) and autocorrelated cross-panel shocks (via a kernel
applied to time-aggregated scores). Specify both `clusters` and `kernel`:

```{r ck}
fit_ck <- ivreg2(
  lwage ~ educ + black + hisp + exper + expersq + married + union,
  data = wagepan, clusters = ~ nr + year, kernel = "bartlett", bw = 3,
  tvar = "year", ivar = "nr", small = TRUE
)
```

The variance decomposition is
$V_{CK} = V_{cluster,i} + V_{kernel,t} - V_{HAC}$, where the third term is
the observation-level HAC variance (the individual-time intersection),
following the two-way structure of Cameron, Gelbach & Miller (2011) and
Thompson (2011). Note that Thompson's specific proposal is this estimator
with the *truncated* kernel, with the bandwidth chosen as the number of
periods after which the common autocorrelated shocks can be ignored
(help-file lines 237--246); the Bartlett variant shown here is a documented
generalization rather than Thompson's own estimator.

### When to use which

| VCE | Cross-sectional heteroskedasticity | Within-panel autocorrelation | Cross-panel dependence |
|-----|-------------------------------------|-------------------------------|------------------------|
| Kiefer | No* | Yes | No |
| Driscoll-Kraay | Yes | Yes (kernel) | Yes |
| Cluster+kernel | Yes (clustering) | Yes (clustering) | Yes (kernel) |
| Two-way cluster | Yes | Yes (within groups) | Yes (within groups) |

*Kiefer assumes conditional homoskedasticity across all observations (both
across individuals and across time), not just across individuals.

## Two-step efficient GMM

### When GMM helps

Two-step efficient GMM (`method = "gmm2s"`) uses an optimal weighting matrix
to exploit overidentifying restrictions more efficiently than 2SLS under
non-IID errors. When exactly identified, GMM2S and 2SLS produce identical
coefficients.

We follow the worked example of Baum, Schaffer & Stillman (2007, p. 476): a
quarterly U.S. Phillips-curve equation from Stock & Watson's textbook
dataset, with the change in inflation regressed on the unemployment rate,
instrumented by lags of GDP growth, the T-bill rate, the exchange rate, and
the T-bond rate. The errors are serially correlated, so the GMM weighting
matrix is HAC (Bartlett kernel, bandwidth 5). First the 2SLS baseline:

```{r sw-2sls}
fit_sw_2sls <- ivreg2(
  dinf ~ 1 | UR | ggdp_2 + TBILL_1 + ER_1 + TBON_1,
  data = stockwatson
)
tidy(fit_sw_2sls)[tidy(fit_sw_2sls)$term == "UR", ]
```

Then HAC two-step GMM:

```{r gmm2s}
fit_gmm <- ivreg2(
  dinf ~ 1 | UR | ggdp_2 + TBILL_1 + ER_1 + TBON_1,
  data = stockwatson, method = "gmm2s",
  vcov = "robust", kernel = "bartlett", bw = 5, tvar = "date"
)
summary(fit_gmm)
```

The first step estimates the model by 2SLS and computes the optimal weighting
matrix from the residuals and the specified VCE type. The second step
re-estimates using this optimal weighting. The Hansen J statistic tests the
overidentifying restrictions. Replicating BSS's narrative: the iid 2SLS
estimate of the unemployment coefficient is
`r sprintf("%.3f", coef(fit_sw_2sls)["UR"])` and clearly significant, while the
HAC-GMM estimate drops to `r sprintf("%.3f", coef(fit_gmm)["UR"])`
(p = `r round(tidy(fit_gmm)$p.value[tidy(fit_gmm)$term == "UR"], 3)`) ---
accounting for the serial correlation removes the apparent significance.

Two finite-sample cautions before reaching for GMM2S by default. The optimal
weighting matrix S is a function of fourth moments of the data, and
estimating fourth moments well can require very large samples; the practical
consequence is that efficient-GMM Wald tests tend to over-reject in finite
samples (Baum, Schaffer & Stillman, 2003, Section 2.7, citing Hayashi 2000,
p. 215). Performance also degrades as the instrument count grows relative to
the sample (Baum, Schaffer & Stillman, 2007, p. 488). In small samples with
modest overidentification, robust 2SLS may be preferable.

### GMM2S with robust/cluster VCE

GMM2S works with all VCE types --- here heteroskedasticity-robust (no
kernel) on the Card schooling data:

```{r gmm2s-robust}
fit_gmm_robust <- ivreg2(
  lwage ~ exper + expersq + black + smsa + south + smsa66 +
    reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669 |
    educ | nearc2 + nearc4,
  data = card, method = "gmm2s", vcov = "robust", small = TRUE
)
tidy(fit_gmm_robust)[tidy(fit_gmm_robust)$term == "educ", ]
```

## CUE estimation

### What CUE does

The Continuously-Updated Estimator (CUE) jointly optimizes the GMM objective
over both the parameter vector and the weighting matrix. CUE is the GMM
generalization of LIML to non-IID errors: with the
conditional-homoskedasticity (iid) weighting, CUE and LIML with
`coviv = TRUE` are mathematically identical, and the computed results agree
up to optimizer tolerance (CUE is found by numerical optimization).

Continuing the Stock-Watson equation, this replicates Baum, Schaffer &
Stillman (2007, Section 4.3, p. 480):

```{r cue}
fit_cue <- ivreg2(
  dinf ~ 1 | UR | ggdp_2 + TBILL_1 + ER_1 + TBON_1,
  data = stockwatson, method = "cue",
  vcov = "robust", kernel = "bartlett", bw = 5, tvar = "date"
)
summary(fit_cue)
```

This completes BSS's narrative arc for the unemployment coefficient: 2SLS
`r sprintf("%.3f", coef(fit_sw_2sls)["UR"])` (significant) → HAC GMM2S
`r sprintf("%.3f", coef(fit_gmm)["UR"])`
(p = `r round(tidy(fit_gmm)$p.value[tidy(fit_gmm)$term == "UR"], 3)`) → CUE
`r sprintf("%.3f", coef(fit_cue)["UR"])`
(p = `r round(tidy(fit_cue)$p.value[tidy(fit_cue)$term == "UR"], 3)`), with
the Hansen J statistic at `r round(fit_cue$diagnostics$overid$stat, 3)`
(p = `r round(fit_cue$diagnostics$overid$p, 4)`).

### Practical caveats

Like LIML, CUE has no finite moments of any order, regardless of the degree
of overidentification, and its finite-sample distribution can have very
heavy tails (Hansen, Heaton & Yaron, 1996). Its objective function can have
multiple local optima, and BSS warn that CUE estimation "can be slow and
problematic when the number of parameters to be estimated is substantial"
(Baum, Schaffer & Stillman, 2007, p. 479). `ivreg2r` uses BFGS optimization
followed by Nelder-Mead refinement, starting from the 2SLS estimates under
iid weighting and from the two-step efficient GMM estimates under a
robust/cluster/kernel VCE (matching Stata's documented behavior). In rare
cases the optimizer may not converge or may find a different basin than
Stata's Newton-Raphson. The `b0` argument (below) can help diagnose
convergence issues.

### Evaluating the CUE objective at a point

Use `b0` to evaluate the CUE objective function at a specific parameter
vector without optimization. The canonical use is constructing the
Stock-Wright (2000) S statistic by hand: the help file (lines 1425--1449)
evaluates the CUE objective on the Mroz data at a point that sets the
endogenous coefficient (`educ`) to zero and the exogenous coefficients to
their OLS values --- fixing the exogenous coefficients at OLS is equivalent
to partialling them out, so the resulting objective value *is* the
Stock-Wright S statistic for H0: the `educ` coefficient is zero and the
orthogonality conditions are valid. Stata displays it with `di e(j)`:

```{r b0}
mroz_work <- subset(mroz, inlf == 1)
ols_fit <- lm(lwage ~ exper + expersq, data = mroz_work)
b0 <- c(educ = 0, coef(ols_fit))

fit_b0 <- ivreg2(
  lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
  data = mroz_work, vcov = "robust", b0 = b0
)
fit_b0$diagnostics$overid$stat  # the objective value, Stata's e(j)
```

The identity is easy to confirm against the Stock-Wright S statistic that
the regular (optimized) fit of the same model reports automatically:

```{r b0-sw-identity}
fit_reg <- ivreg2(
  lwage ~ exper + expersq | educ | age + kidslt6 + kidsge6,
  data = mroz_work, vcov = "robust"
)
c(b0_objective  = fit_b0$diagnostics$overid$stat,
  stock_wright_S = fit_reg$diagnostics$stock_wright$stat)
```

The correct reference distribution is chi-squared with
L1 = `r length(fit_b0$instruments)` (number of excluded instruments)
degrees of freedom, giving
p = `r round(fit_reg$diagnostics$stock_wright$p, 4)`. One caution: the
p-value stored alongside the raw objective in `fit_b0$diagnostics$overid$p`
is calibrated against L − K overidentification degrees of freedom, which is
*not* the right reference distribution for this use --- display the raw
value, as Stata's `di e(j)` does, and take the S-statistic inference from
the regular fit's summary.

(The canonical `b0` example is cross-sectional; the mechanism is identical
under a kernel-based VCE. Evaluating the objective at a candidate optimum
also remains a useful convergence check for CUE fits.)

## User-supplied W and S matrices

For iterated GMM or custom weighting, you can supply your own weighting
matrix (`wmatrix`, Stata's `wmatrix()`) and/or moment-covariance matrix
(`smatrix`, Stata's `smatrix()`). When `wmatrix` is supplied alone, the VCV
uses a sandwich formula with the user's W as bread and the estimated S as
filling. When both are supplied, S is used directly for the VCV and
J-statistic. When `wmatrix` is used with `method = "gmm2s"`, the user's W is
used as the first-step weighting matrix.

The canonical workflow (help-file lines 1156--1198) re-uses the estimated S
and W from one fit in a subsequent estimation --- in Stata these are posted
on the fitted model as `e(S)` and `e(W)`. `ivreg2r` does not yet export the
estimated S and W matrices on the fitted object, so that round-trip cannot
be demonstrated here; the export is planned for a future release. A matrix
you construct yourself (with the 1/N normalization `smatrix` expects) works
today.

## Partialling out regressors

### FWL projection

The `partial` argument removes specified exogenous regressors from all
variables (dependent, endogenous, excluded instruments) via Frisch-Waugh-Lovell
projection before estimation. This is useful for absorbing high-dimensional
fixed effects without including them as regressors:

```{r partial}
fit_partial <- ivreg2(
  lwage ~ educ + black + hisp + exper + expersq + married + union,
  data = wagepan,
  partial = c("black", "hisp"),
  clusters = ~ nr, small = TRUE
)
```

Coefficients on partialled-out variables are not reported (they are projected
out before estimation), but the remaining coefficients are numerically
identical to including them as regressors. This equivalence holds for 2SLS,
LIML, and two-step GMM estimation. It does not hold for CUE or iterated GMM
beyond two steps.

### Special tokens

- `"_cons"` partials out the constant (equivalent to demeaning all variables).
- `"_all"` partials out all exogenous regressors, leaving only the
  endogenous variables and excluded instruments.

```{r partial-cons}
# Demean all variables (partial out the constant)
fit_demean <- ivreg2(cinf ~ unem, data = phillips, partial = "_cons")
```

### DoF adjustment

By default, partialling adjusts `sdofminus` to account for the degrees of
freedom consumed by the partialled-out variables. Use `nopartialsmall = TRUE`
to suppress this adjustment:

```{r partial-nosmall, eval = FALSE}
fit_np <- ivreg2(cinf ~ unem + unem_1, data = phillips,
                 partial = "_cons", nopartialsmall = TRUE)
```

## Additional options

### Redundancy test

The `redundant` argument tests whether specific excluded instruments are
redundant --- whether, given the other instruments, using them improves the
asymptotic efficiency of the estimation. This uses an LM rank test: the
Anderson canonical correlations statistic under iid errors, or the
Kleibergen-Paap rk statistic with a robust, cluster, or kernel VCE:

```{r redundant}
fit_red <- ivreg2(
  lwage ~ exper + expersq + black + smsa + south + smsa66 +
    reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669 |
    educ | nearc2 + nearc4,
  data = card, vcov = "robust", redundant = c("nearc2")
)
glance(fit_red)[, c("redundancy_stat", "redundancy_p")]
```

The null hypothesis is that the specified instruments are redundant.
Redundancy is about *efficiency*, not identification: excluded instruments
are redundant if using them does not improve the asymptotic efficiency of
the estimation (help-file lines 560--563), conditional on the model being
identified by the remaining instruments. A rejection means the tested
instruments are conditionally relevant --- they reduce the asymptotic
variance of the estimates and are worth keeping *provided their validity is
maintained*; the redundancy test says nothing about whether they satisfy
the exclusion restrictions.

### Centering moment conditions

The `center` option subtracts the mean of the moment conditions before
computing the VCE and the long-run covariance matrix S used in diagnostics.
This can improve the finite-sample properties of standard errors and the
J-test:

```{r center}
fit_cen <- ivreg2(
  lwage ~ exper + expersq + black + smsa + south + smsa66 +
    reg662 + reg663 + reg664 + reg665 + reg666 + reg667 + reg668 + reg669 |
    educ | nearc2 + nearc4,
  data = card, method = "gmm2s", vcov = "robust", center = TRUE
)
```

### PSD corrections

Kernels outside the Bartlett/Parzen/Quadratic-Spectral family do not
guarantee a positive semi-definite estimate of S, the covariance matrix of
the orthogonality conditions that feeds both the VCE and the chi-squared
diagnostics. When S has negative eigenvalues, `psd` applies an eigenvalue
correction to it:

- `psd = "psd0"`: set negative eigenvalues to zero (Politis, 2007).
- `psd = "psda"`: replace negative eigenvalues with their absolute values
  (Stock & Watson, 2008, Remark 8).

Note the example below can never trigger the correction: Driscoll-Kraay with
the default Bartlett kernel is positive semi-definite by construction
(Newey-West, 1987), so `psd` is shown purely for syntax. The option earns
its keep with non-PSD kernels such as the truncated kernel.

```{r psd, eval = FALSE}
fit_psd <- ivreg2(
  lwage ~ educ + exper + expersq + married + union,
  data = wagepan, dkraay = 3, tvar = "year", ivar = "nr",
  psd = "psd0"
)
```

### DoF adjustments

When fixed effects or other controls have been partialled out manually
(outside of `ivreg2()`), the `dofminus` and `sdofminus` arguments adjust the
degrees of freedom. See the Degrees-of-freedom adjustments section of
`vignette("advanced-iv")` for the full exposition, including a self-verifying
Frisch-Waugh-Lovell demonstration of `sdofminus`.

## Estimator equivalences

Several estimators are mathematically equivalent under specific conditions,
which can be useful for verification and understanding:

| Equivalence | Conditions |
|-------------|------------|
| LIML + COVIV = CUE | Conditional-homoskedasticity (iid) weighting; agreement up to optimizer tolerance |
| Driscoll-Kraay = cluster(time) + kernel | Same kernel and bandwidth |
| Kiefer = truncated kernel | Panel data, AC, `bw = T - 1` (all within-panel lags weighted one) |
| LIML = 2SLS | Exactly identified |
| GMM2S = 2SLS | Exactly identified |

## References

- Andrews, D.W.K. (1991). "Heteroskedasticity and Autocorrelation Consistent
  Covariance Matrix Estimation." *Econometrica*, 59(3), 817--858.
- Baum, C.F., Schaffer, M.E., and Stillman, S. (2003). "Instrumental
  Variables and GMM: Estimation and Testing." *Stata Journal*, 3(1), 1--31.
- Baum, C.F., Schaffer, M.E., and Stillman, S. (2007). "Enhanced Routines for
  Instrumental Variables/Generalized Method of Moments Estimation and
  Testing." *Stata Journal*, 7(4), 465--506.
- Cameron, A.C., Gelbach, J.B., and Miller, D.L. (2011). "Robust Inference
  with Multiway Clustering." *Journal of Business & Economic Statistics*,
  29(2), 238--249.
- Driscoll, J.C. and Kraay, A.C. (1998). "Consistent Covariance Matrix
  Estimation with Spatially Dependent Panel Data." *Review of Economics and
  Statistics*, 80(4), 549--560.
- Hansen, L.P. (1982). "Large Sample Properties of Generalized Method of
  Moments Estimators." *Econometrica*, 50(4), 1029--1054.
- Hansen, L.P., Heaton, J., and Yaron, A. (1996). "Finite-Sample Properties
  of Some Alternative GMM Estimators." *Journal of Business & Economic
  Statistics*, 14(3), 262--280.
- Hayashi, F. (2000). *Econometrics*. Princeton University Press.
- Kiefer, N.M. (1980). "Estimation of Fixed Effect Models for Time Series of
  Cross-Sections with Arbitrary Intertemporal Covariance." *Journal of
  Econometrics*, 14(2), 195--202.
- Kleibergen, F. and Paap, R. (2006). "Generalized Reduced Rank Tests Using
  the Singular Value Decomposition." *Journal of Econometrics*, 133(1),
  97--126.
- Mroz, T.A. (1987). "The Sensitivity of an Empirical Model of Married
  Women's Hours of Work to Economic and Statistical Assumptions."
  *Econometrica*, 55(4), 765--799.
- Newey, W.K. and West, K.D. (1987). "A Simple, Positive Semi-Definite,
  Heteroskedasticity and Autocorrelation Consistent Covariance Matrix."
  *Econometrica*, 55(3), 703--708.
- Newey, W.K. and West, K.D. (1994). "Automatic Lag Selection in Covariance
  Matrix Estimation." *Review of Economic Studies*, 61(4), 631--653.
- Politis, D.N. (2007). "Higher-Order Accurate, Positive Semi-Definite
  Estimation of Large-Sample Covariance and Spectral Density Matrices."
  Working paper, UC San Diego; published (2011) in *Econometric Theory*,
  27(4), 703--744.
- Stock, J.H. and Watson, M.W. (2008). "Heteroskedasticity-Robust Standard
  Errors for Fixed Effects Panel Data Regression." *Econometrica*, 76(1),
  155--174.
- Stock, J.H. and Wright, J.H. (2000). "GMM with Weak Identification."
  *Econometrica*, 68(5), 1055--1096.
- Stock, J.H. and Yogo, M. (2005). "Testing for Weak Instruments in Linear IV
  Regression." In D.W.K. Andrews and J.H. Stock (Eds.), *Identification and
  Inference for Econometric Models: Essays in Honor of Thomas Rothenberg*.
  Cambridge University Press.
- Thompson, S.B. (2011). "Simple Formulas for Standard Errors that Cluster by
  Both Firm and Time." *Journal of Financial Economics*, 99(1), 1--10.
