---
title: "Time Series IV, Panel VCE, and GMM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Time Series IV, Panel VCE, and GMM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

When errors are autocorrelated, heteroskedastic across time, or correlated
across individuals in a panel, the classical and HC-robust standard errors
from `vignette("introduction")` are inconsistent. This vignette covers the
tools `ivreg2r` provides for these settings:

- **HAC and AC standard errors** for time-series data
- **Panel VCE options**: Kiefer, Driscoll-Kraay, and cluster+kernel
- **Two-step efficient GMM** for overidentified models with non-IID errors
- **CUE estimation** as the GMM generalization of LIML
- **User-supplied W and S matrices** for iterated GMM and custom weighting
- **Partialling out regressors** for high-dimensional fixed effects
- **Additional options**: redundancy tests, moment centering, PSD corrections

We use two datasets: the Phillips (1958) curve data (pure time series) and the
Wooldridge wage panel.

```{r setup}
library(ivreg2r)
data(phillips)
data(wagepan)
```

## HAC and AC standard errors

### The Phillips curve

The Phillips dataset contains 49 annual U.S. observations (1948--1996) on
inflation and unemployment. A simple expectations-augmented Phillips curve
regresses the change in inflation on unemployment:

```{r phillips-ols}
phil <- na.omit(phillips)
fit_iid <- ivreg2(cinf ~ unem, data = phil)
```

With time-series data, the errors are likely autocorrelated, making classical
SEs inconsistent. HAC (heteroskedasticity and autocorrelation consistent)
standard errors account for both.

### Specifying kernel-based standard errors

Use `kernel`, `bw` (bandwidth), and `tvar` to request kernel-based inference.
The type of VCE depends on the `vcov` setting: with `vcov = "robust"` (or
when heteroskedasticity-robust VCE is already active), you get **HAC**
(heteroskedasticity and autocorrelation consistent) standard errors. With the
default `vcov = "iid"`, specifying a kernel gives **AC** (autocorrelation
consistent) standard errors, which assume homoskedasticity but allow for
serial correlation. This matches Stata's `ivreg2`, where `bw`/`kernel` alone
produces AC and combining `robust` with `bw` produces HAC.

```{r hac-bartlett}
fit_hac <- ivreg2(cinf ~ unem, data = phil,
                  vcov = "robust", kernel = "bartlett", bw = 3, tvar = "year")
summary(fit_hac)
```

The Bartlett kernel applies linearly declining weights to autocovariances up
to lag `bw - 1`. This is the Newey-West (1987) estimator and is the most
common choice in applied work.

### Choosing a kernel

`ivreg2r` supports eight kernels, matching Stata's `ivreg2`:

| Kernel | Weights | Guaranteed PSD? | Typical use |
|--------|---------|-----------------|-------------|
| Bartlett | Linear decline | Yes | Default choice; Newey-West |
| Parzen | Piecewise cubic | Yes | Smoother than Bartlett |
| Quadratic Spectral | $\sin/\cos$ based | No | Optimal MSE rate |
| Truncated | Uniform up to lag | No | All lags equally weighted; Kiefer equivalence |
| Tukey-Hanning | Cosine taper | No | Standard spectral window |
| Tukey-Hamming | Cosine taper | No | Variant of Hanning |
| Daniell | $\sin(x)/x$ | No | Spectral analysis |
| Tent | Triangular | No | Simple triangular window |

Only Bartlett and Parzen guarantee a positive semi-definite covariance estimate.
For other kernels, use the `psd` option if negative eigenvalues arise.

```{r hac-kernels}
fit_qs <- ivreg2(cinf ~ unem, data = phil,
                 kernel = "qs", bw = 3, tvar = "year")
```

### Automatic bandwidth

Use `bw = "auto"` for the Newey-West (1994) plug-in bandwidth selector, which
chooses the bandwidth from the data. This is available for Bartlett, Parzen,
and Quadratic Spectral kernels:

```{r hac-auto}
fit_auto <- ivreg2(cinf ~ unem, data = phil,
                   kernel = "bartlett", bw = "auto", tvar = "year")
fit_auto$bw  # selected bandwidth
```

### HAC vs AC

AC (autocorrelation consistent) standard errors assume homoskedasticity but
allow for autocorrelation. Use `vcov = "AC"` explicitly:

```{r ac}
fit_ac <- ivreg2(cinf ~ unem, data = phil,
                 vcov = "AC", kernel = "bartlett", bw = 3, tvar = "year")
```

HAC allows both heteroskedasticity and autocorrelation; AC assumes
homoskedastic errors with autocorrelation. AC is more efficient when
homoskedasticity holds, at the cost of inconsistency if it does not.

### HAC with IV

HAC standard errors work with all estimation methods. Here we instrument
unemployment with its second lag:

```{r hac-iv}
fit_iv_hac <- ivreg2(
  cinf ~ unem_1 | unem | unem_2,
  data = phil, kernel = "bartlett", bw = 3, tvar = "year"
)
summary(fit_iv_hac)
```

All diagnostics (overidentification, identification, endogeneity) automatically
adapt to the HAC VCE.

## Panel VCE options

Panel data introduces richer correlation structures. Beyond one-way and
two-way clustering (covered in `vignette("introduction")` and
`vignette("advanced-iv")`), `ivreg2r` provides three kernel-based panel VCE
options.

### Kiefer VCE

Kiefer (1980) VCE assumes conditionally homoskedastic errors (across all
observations --- both across individuals and across time) but allows arbitrary
within-panel serial correlation. It is equivalent to AC VCE with a truncated
kernel where the bandwidth equals T (all within-panel lags equally weighted):

```{r kiefer}
fit_kiefer <- ivreg2(
  lwage ~ educ + black + hisp + exper + expersq + married + union,
  data = wagepan, kiefer = TRUE, tvar = "year", ivar = "nr",
  small = TRUE
)
```

Use Kiefer when you believe the cross-sectional units are homogeneous but
want to allow for arbitrary within-unit serial correlation.

### Driscoll-Kraay VCE

Driscoll and Kraay (1998) VCE handles cross-sectional dependence *and*
persistence. It clusters on the time variable and applies a kernel to scores
summed within each time period:

```{r dk}
fit_dk <- ivreg2(
  lwage ~ educ + black + hisp + exper + expersq + married + union,
  data = wagepan, dkraay = 3, tvar = "year", ivar = "nr",
  small = TRUE
)
```

The `dkraay` argument specifies the bandwidth (here, Bartlett with 3 lags).
DK is robust to arbitrary cross-sectional dependence --- useful when common
shocks (business cycles, policy changes) affect all individuals
simultaneously. DK is equivalent to clustering on the time variable with a
kernel (Bartlett by default; can be overridden with the `kernel` argument).

**Important:** Driscoll-Kraay is a large-T estimator --- the panel should have
a reasonably long time-series dimension for the asymptotics to apply. It is
not appropriate for panels with very few time periods.

### Cluster + kernel (Thompson)

Thompson (2011) two-way cluster+kernel VCE accounts for both within-panel
correlation (via clustering on `ivar`) and cross-panel persistence (via a
kernel applied to time-aggregated scores). Specify both `clusters` and
`kernel`:

```{r ck}
fit_ck <- ivreg2(
  lwage ~ educ + black + hisp + exper + expersq + married + union,
  data = wagepan, clusters = ~ nr + year, kernel = "bartlett", bw = 3,
  tvar = "year", ivar = "nr", small = TRUE
)
```

This uses the Cameron-Gelbach-Miller decomposition:
$V_{CK} = V_{cluster,i} + V_{kernel,t} - V_{HAC}$, where the third term is the
observation-level HAC variance (the individual-time intersection).

### When to use which

| VCE | Cross-sectional heteroskedasticity | Within-panel autocorrelation | Cross-panel dependence |
|-----|-------------------------------------|-------------------------------|------------------------|
| Kiefer | No* | Yes | No |
| Driscoll-Kraay | Yes | Yes (kernel) | Yes |
| Cluster+kernel | Yes (clustering) | Yes (clustering) | Yes (kernel) |
| Two-way cluster | Yes | Yes (within groups) | Yes (within groups) |

*Kiefer assumes conditional homoskedasticity across all observations (both
across individuals and across time), not just across individuals.

## Two-step efficient GMM

### When GMM helps

Two-step efficient GMM (`method = "gmm2s"`) uses an optimal weighting matrix
to exploit overidentifying restrictions more efficiently than 2SLS under
non-IID errors. When exactly identified, GMM2S and 2SLS produce identical
coefficients.

```{r gmm2s}
fit_gmm <- ivreg2(
  cinf ~ unem_1 | unem | unem_2 + inf_2,
  data = phil, method = "gmm2s",
  kernel = "bartlett", bw = 3, tvar = "year"
)
summary(fit_gmm)
```

The first step estimates the model by 2SLS and computes the optimal weighting
matrix from the residuals and the specified VCE type. The second step
re-estimates using this optimal weighting. The Hansen J statistic tests the
overidentifying restrictions.

### GMM2S with robust/cluster VCE

GMM2S works with all VCE types:

```{r gmm2s-robust}
fit_gmm_robust <- ivreg2(
  lwage ~ exper + expersq + black + south + smsa + married |
    educ | nearc2 + nearc4,
  data = card, method = "gmm2s", vcov = "robust", small = TRUE
)
```

In practice, efficiency gains of GMM2S over 2SLS are modest for
nearly-identified models (few overidentifying restrictions). The gains are
larger when there are many instruments relative to endogenous variables.

## CUE estimation

### What CUE does

The Continuously-Updated Estimator (CUE) jointly optimizes the GMM objective
over both the parameter vector and the weighting matrix. CUE is the GMM
generalization of LIML to non-IID errors: under iid errors (no
robust/cluster/kernel VCE), CUE and LIML with `coviv = TRUE` produce identical
results to machine precision.

```{r cue}
fit_cue <- ivreg2(
  cinf ~ unem_1 | unem | unem_2 + inf_2,
  data = phil, method = "cue",
  kernel = "bartlett", bw = 3, tvar = "year"
)
summary(fit_cue)
```

### Practical caveats

Like LIML, CUE may lack finite moments, particularly when the degree of
overidentification is small. Its objective function can have
multiple local optima. `ivreg2r` uses BFGS optimization with 2SLS starting
values, followed by Nelder-Mead refinement. In rare cases the optimizer may
not converge or may find a different basin than Stata's Newton-Raphson. The
`b0` argument (see below) can help diagnose convergence issues.

### Evaluating the CUE objective at a point

Use `b0` to evaluate the CUE objective function at a specific parameter
vector without optimization. This is useful for GMM-distance tests or
verifying convergence:

```{r b0, eval = FALSE}
# Evaluate CUE J-statistic at 2SLS estimates
b_2sls <- coef(ivreg2(cinf ~ unem_1 | unem | unem_2 + inf_2,
                       data = phil))
fit_b0 <- ivreg2(
  cinf ~ unem_1 | unem | unem_2 + inf_2,
  data = phil, method = "cue", b0 = b_2sls,
  kernel = "bartlett", bw = 3, tvar = "year"
)
fit_b0$diagnostics$overid  # J at 2SLS point
```

## User-supplied W and S matrices

For iterated GMM or custom weighting, you can supply your own weighting
matrix (`wmatrix`) and/or long-run covariance matrix (`smatrix`):

```{r wmatrix, eval = FALSE}
# Use identity weighting matrix (heterogeneous GMM)
n_inst <- 10  # number of instruments including exogenous
W <- diag(n_inst)
fit_w <- ivreg2(
  lwage ~ exper + expersq | educ | nearc2 + nearc4,
  data = card, wmatrix = W, vcov = "robust"
)

# Supply S matrix for VCV and J computation
S <- fit_w$omega  # use the Omega from a previous estimation
fit_ws <- ivreg2(
  lwage ~ exper + expersq | educ | nearc2 + nearc4,
  data = card, wmatrix = W, smatrix = S, vcov = "robust"
)
```

When `wmatrix` is supplied alone, the VCV uses a sandwich formula with the
user's W as bread and the estimated S as filling. When both are supplied, S
is used directly for the VCV and J-statistic. When `wmatrix` is used with
`method = "gmm2s"`, the user's W is used as the first-step weighting matrix.

## Partialling out regressors

### FWL projection

The `partial` argument removes specified exogenous regressors from all
variables (dependent, endogenous, excluded instruments) via Frisch-Waugh-Lovell
projection before estimation. This is useful for absorbing high-dimensional
fixed effects without including them as regressors:

```{r partial}
fit_partial <- ivreg2(
  lwage ~ educ + black + hisp + exper + expersq + married + union,
  data = wagepan,
  partial = c("black", "hisp"),
  clusters = ~ nr, small = TRUE
)
```

Coefficients on partialled-out variables are not reported (they are projected
out before estimation), but the remaining coefficients are numerically
identical to including them as regressors. This equivalence holds for 2SLS,
LIML, and two-step GMM estimation. It does not hold for CUE or iterated GMM
beyond two steps.

### Special tokens

- `"_cons"` partials out the constant (equivalent to demeaning all variables).
- `"_all"` partials out all exogenous regressors, leaving only the
  endogenous variables and excluded instruments.

```{r partial-cons}
# Demean all variables (partial out the constant)
fit_demean <- ivreg2(cinf ~ unem, data = phil, partial = "_cons")
```

### DoF adjustment

By default, partialling adjusts `sdofminus` to account for the degrees of
freedom consumed by the partialled-out variables. Use `nopartialsmall = TRUE`
to suppress this adjustment:

```{r partial-nosmall, eval = FALSE}
fit_np <- ivreg2(cinf ~ unem + unem_1, data = phil,
                 partial = "_cons", nopartialsmall = TRUE)
```

## Additional options

### Redundancy test

The `redundant` argument tests whether specific excluded instruments are
redundant (add no identifying information beyond the other instruments). This
uses an LM rank test: the Anderson canonical correlations statistic under iid
errors, or the Kleibergen-Paap rk statistic with a robust, cluster, or kernel
VCE:

```{r redundant}
fit_red <- ivreg2(
  lwage ~ exper + expersq + black + south + smsa + married |
    educ | nearc2 + nearc4,
  data = card, vcov = "robust", redundant = c("nearc2")
)
glance(fit_red)[, c("redundancy_stat", "redundancy_p")]
```

The null hypothesis is that the specified instruments are redundant. Rejection
means they contribute to identification.

### Centering moment conditions

The `center` option subtracts the mean of the moment conditions before
computing the VCE and the long-run covariance matrix S used in diagnostics.
This can improve the finite-sample properties of standard errors and the
J-test:

```{r center}
fit_cen <- ivreg2(
  lwage ~ exper + expersq + black + south + smsa + married |
    educ | nearc2 + nearc4,
  data = card, method = "gmm2s", vcov = "robust", center = TRUE
)
```

### PSD corrections

When the estimated VCV matrix has negative eigenvalues (possible with
cluster-kernel or DK VCE), `psd` applies a correction:

- `psd = "psd0"`: set negative eigenvalues to zero.
- `psd = "psda"`: replace negative eigenvalues with their absolute values.

```{r psd, eval = FALSE}
fit_psd <- ivreg2(
  lwage ~ educ + exper + expersq + married + union,
  data = wagepan, dkraay = 3, tvar = "year", ivar = "nr",
  psd = "psd0"
)
```

### DoF adjustments

When fixed effects or other controls have been partialled out manually (outside
of `ivreg2()`), use `dofminus` and `sdofminus` to adjust degrees of freedom:

- `dofminus`: subtracted from N in large-sample variance formulas.
- `sdofminus`: subtracted from residual df alongside K in small-sample
  corrections.

```{r dofminus}
fit_dof <- ivreg2(cinf ~ unem, data = phil,
                  kernel = "bartlett", bw = 3, tvar = "year",
                  dofminus = 2, small = TRUE)
```

## Estimator equivalences

Several estimators are mathematically equivalent under specific conditions,
which can be useful for verification and understanding:

| Equivalence | Conditions |
|-------------|------------|
| LIML + COVIV = CUE | IID errors (no robust/cluster/kernel VCE) |
| Driscoll-Kraay = cluster(time) + kernel | Same kernel and bandwidth |
| Kiefer = truncated kernel at full bandwidth | Panel data, `bw = T` |
| LIML = 2SLS | Exactly identified |
| GMM2S = 2SLS | Exactly identified |

## References

- Driscoll, J.C. and Kraay, A.C. (1998). "Consistent Covariance Matrix
  Estimation with Spatially Dependent Panel Data." *Review of Economics and
  Statistics*, 80(4), 549--560.
- Hansen, L.P. (1982). "Large Sample Properties of Generalized Method of
  Moments Estimators." *Econometrica*, 50(4), 1029--1054.
- Hansen, L.P., Heaton, J., and Yaron, A. (1996). "Finite-Sample Properties
  of Some Alternative GMM Estimators." *Journal of Business & Economic
  Statistics*, 14(3), 262--280.
- Kiefer, N.M. (1980). "Estimation of Fixed Effect Models for Time Series of
  Cross-Sections with Arbitrary Intertemporal Covariance." *Journal of
  Econometrics*, 14(2), 195--202.
- Newey, W.K. and West, K.D. (1987). "A Simple, Positive Semi-Definite,
  Heteroskedasticity and Autocorrelation Consistent Covariance Matrix."
  *Econometrica*, 55(3), 703--708.
- Newey, W.K. and West, K.D. (1994). "Automatic Lag Selection in Covariance
  Matrix Estimation." *Review of Economic Studies*, 61(4), 631--653.
- Stock, J.H. and Yogo, M. (2005). "Testing for Weak Instruments in Linear IV
  Regression." In D.W.K. Andrews and J.H. Stock (Eds.), *Identification and
  Inference for Econometric Models: Essays in Honor of Thomas Rothenberg*.
  Cambridge University Press.
- Thompson, S.B. (2011). "Simple Formulas for Standard Errors that Cluster by
  Both Firm and Time." *Journal of Financial Economics*, 99(1), 1--10.
- Wooldridge, J.M. (2010). *Econometric Analysis of Cross Section and Panel
  Data*, 2nd ed. MIT Press.
